---
output: html_document
editor_options: 
  chunk_output_type: console
---

## Student t-test

We perform a paired two sided t-test on each sequence to see if there is the mean expression of the ALT and REF sequence are different.

Before performing the t-test, we normalized count using TMM method implemented in edgeR package.

For example, we can compute for rs79227574-ALT-Middle-1 and rs79227574-REF-Middle-1 if there mean expression in both replicate are different.

```{r}
# Function to run two sided t test on two vectors. Return NA is the results can't be run (i.e Ratio for condition have null deviation)
# this function take a matrix as input and an index or rowname for compared contidions. 
student_test_rs_vs_not_rs <- function(input_matrix, cond1, cond2){
  try_student_results_pval <- try(t.test(x = input_matrix[cond1, ], 
                                         y = input_matrix[cond2, ],
                                         paired = T, alternative = "two.sided")$p.value, silent = T)
  try_student_results_estimate <- try(t.test(x = input_matrix[cond1, ], 
                                             y = input_matrix[cond2, ],
                                             paired = T, alternative = "two.sided")$estimate, silent = T)
  if(grepl(pattern = "Error", x = try_student_results_pval) | grepl(pattern = "Error", x = try_student_results_estimate)){
    student_results_pval <- NA
    student_results_estimate <- NA
  }else{
    student_results_pval <- try_student_results_pval
    student_results_estimate <- try_student_results_estimate
  }
  return(list(pval = student_results_pval, estimate = student_results_estimate))
}
```

```{r}
# Normalize data using TMM method
scale.factors <- calcNormFactors(count_matrix,lib.size=NULL, method = "TMM")
lib.size <- colSums(count_matrix)
norm.data <- t(t(count_matrix)/(scale.factors*lib.size))

cDNA_matrix_norm <- count_matrix[,grep("cDNA", colnames(count_matrix))]
gDNA_matrix_norm <- count_matrix[,grep("ADNg", colnames(count_matrix))]
# Prepare data for t-test on ALT/REF
ratio <- (cDNA_matrix_norm+1)/(gDNA_matrix_norm+1)
ratio_matrix <- log2(ratio)
colnames(ratio_matrix) <- gsub(pattern = "_cDNA|_ADNg", replacement = "", x = colnames(ratio_matrix))

# loop on condition
condition <- c("101_replicat", "HSO_replicat")

alt_rs_name <- rownames(ratio_matrix[grep("ALT", rownames(ratio_matrix)),])
# for(cond in condition){
#   cond_ratio_matrix <- ratio_matrix[,grep(pattern = cond, x = colnames(ratio_matrix))]
#   for(rs_name in alt_rs_name){
#     rs_name_unique <- gsub(pattern = "ALT-", replacement = "", x = rs_name)
#     student_results <- t.test(x = cond_ratio_matrix[rs_name, ], 
#                               y = cond_ratio_matrix[gsub(pattern = "ALT", replacement = "REF", x = rs_name), ],
#                               paired = T, alternative = "two.sided")
#     ttest_results_matrix[rs_name_unique, "student_t_value"] <- student_results$statistic
#     ttest_results_matrix[rs_name_unique, "student_p_value"] <- student_results$p.value
#   }
# }

# ttest_results_matrix <- matrix(data = NA, nrow = dim(count_matrix[grep("ALT", rownames(count_matrix)),]), ncol = 2)
# rownames(ttest_results_matrix) <- gsub(pattern = "ALT-", replacement = "",
#                                        rownames(count_matrix[grep("ALT", rownames(count_matrix)),]))
# colnames(ttest_results_matrix) <- c("student_t_value", "student_p_value")

ttest_results_matrix <- matrix(data = NA, nrow = nrow(count_matrix), ncol = 4)
rownames(ttest_results_matrix) <- rownames(count_matrix)
colnames(ttest_results_matrix) <- c("HSO_student_estimate_value", "HSO_student_p_value", "101_student_estimate_value", "101_student_p_value")

for(Alt in 1:3){
  
  grep_ref = paste0("-", Alt)
  
  Alt_names_loop <- Alt_names[grep(grep_ref, Alt_names)]
  Ref_names_loop <- gsub(pattern = grep_ref, replacement = "-1",
                         x = gsub(pattern = "ALT", replacement = "REF", x = Alt_names_loop))
  
  alt_rs_name <- rownames(ratio_matrix[grep("ALT", rownames(ratio_matrix)),])
  for(cond in condition){
    cond_ratio_matrix <- ratio_matrix[,grep(pattern = cond, x = colnames(ratio_matrix))]
    for(rs_name_index in 1:length(Alt_names_loop)){
      rs_name <- gsub(pattern = "-ALT", replacement = "", Alt_names_loop[rs_name_index])
      alt_name_loop <- Alt_names_loop[rs_name_index]
      ref_name_loop <- Ref_names_loop[rs_name_index]
      student_test_results <- student_test_rs_vs_not_rs(input_matrix = cond_ratio_matrix, cond1 = alt_name_loop, cond2 = ref_name_loop)
      ttest_results_matrix[alt_name_loop, paste0(gsub("_replicat", "_", cond), "student_estimate_value")] <- student_test_results$estimate
      ttest_results_matrix[alt_name_loop, paste0(gsub("_replicat", "_", cond), "student_p_value")] <- student_test_results$pval
      ttest_results_matrix[ref_name_loop, paste0(gsub("_replicat", "_", cond), "student_estimate_value")] <- student_test_results$estimate
      ttest_results_matrix[ref_name_loop, paste0(gsub("_replicat", "_", cond), "student_p_value")] <- student_test_results$pval
    }
  }
}
```

```{r}
ttest_results_matrix_for_plot <- as.data.frame(cbind(ratio_matrix, ttest_results_matrix))

ttest_results_matrix_for_plot[is.na(ttest_results_matrix_for_plot)] <- 1

adjusted_pval <- apply(ttest_results_matrix_for_plot[, grep("p_value", colnames(ttest_results_matrix_for_plot))], 2, mt.rawp2adjp, proc="BH")
ttest_results_matrix_for_plot$HSO_student_adj_p_value <- adjusted_pval$HSO_student_p_value$adjp[order(adjusted_pval$HSO_student_p_value$index),][,2]
ttest_results_matrix_for_plot$`101_student_adj_p_value` <- adjusted_pval$`101_student_p_value`$adjp[order(adjusted_pval$`101_student_p_value`$index),][,2]
ttest_results_matrix_for_plot$HSO_student_qvalue <- qvalue(ttest_results_matrix_for_plot$HSO_student_p_value)$qvalues
ttest_results_matrix_for_plot$HSO_qval_pvalue_FDR_5 <- as.factor(qvalue(ttest_results_matrix_for_plot$HSO_student_p_value, fdr.level = 0.05)$significant)
ttest_results_matrix_for_plot$HSO_qval_pvalue_FDR_10 <- as.factor(qvalue(ttest_results_matrix_for_plot$HSO_student_p_value, fdr.level = 0.10)$significant)
ttest_results_matrix_for_plot$`101_student_qvalue` <- qvalue(ttest_results_matrix_for_plot$`101_student_p_value`)$qvalues
ttest_results_matrix_for_plot$`101_qval_adj_pvalue_FDR_5` <- as.factor(qvalue(ttest_results_matrix_for_plot$`101_student_p_value`, fdr.level = 0.05)$significant)
ttest_results_matrix_for_plot$`101_qval_adj_pvalue_FDR_10` <- as.factor(qvalue(ttest_results_matrix_for_plot$`101_student_p_value`, fdr.level = 0.10)$significant)

ttest_results_matrix_for_plot$location <-  gsub(pattern = "rs[0-9]*-|(ALT|REF)-|-[123]", replacement = "", x = rownames(ttest_results_matrix_for_plot))
ttest_results_matrix_for_plot$allele <-  gsub(pattern = "rs[0-9]*-|-(Left|Middle|Right)|-123", replacement = "", x = rownames(ttest_results_matrix_for_plot))
```

### Results for 101 sample

<!-- # ```{r} -->
<!-- # ggplot(ttest_results_matrix_for_plot, aes(x = `101_student_estimate_value`, y = -log10(`101_student_p_value`), color = location)) +  -->
<!-- #   geom_point() + facet_wrap(~ location) + -->
<!-- #   ggtitle("Violin plot of the p-values of the Student test and the estimated difference of means,\nsplitted by location for sample 101") -->
<!-- # ggplot(ttest_results_matrix_for_plot, aes(x = `101_student_estimate_value`, y = -log10(`101_student_p_value`), color = allele)) +  -->
<!-- #   geom_point() + facet_wrap(~ allele) + -->
<!-- #   ggtitle("Violin plot of the p-values of the Student test and the estimated difference of means,\nsplitted by allele for sample 101") -->
<!-- # ``` -->
<!-- #  -->
<!-- # The violin plots above show the distribution of Student test's p-values among each sequence for sample 101. I draw those plots to try to see if there is a major difference between location and/or allele. We can't detect a real difference here. -->

```{r}
hist(ttest_results_matrix_for_plot$`101_student_p_value`, main = "Distribution of raw p-value of the Student test", ylab = "Student's test p-values")
hist(ttest_results_matrix_for_plot$`101_student_adj_p_value`, main = "Distribution of adjusted p-value of the Student test", ylab = "Student's test adjusted p-values")

signif_p_val_nb_101 <- sum(ttest_results_matrix_for_plot$`101_student_p_value` < 0.05)
```

When looking at the adjusted p-values (by the Benjamini Hochberg method), none of them are significant (i.e lower than 0.05).

<!-- #### List of significant sequences for the Student test (i.e p-value lower than 0.05) for sample 101 -->

<!-- There are `r print(signif_p_val_nb_101)` sequences with a significant p-value for the Student test. The list can be obtained using the table "Table_Student_tests_chapter_2.6.csv" or below as an interactive table -->

<!-- ```{r} -->
<!-- datatable(data.frame(sequence_name = rownames(ttest_results_matrix_for_plot[ttest_results_matrix_for_plot$`101_student_p_value`<0.05,]))) -->
<!-- ``` -->

<!-- The following violin plot shows the number of sequences that are control sequences and also have a significant p-value for the Student test. -->

<!-- ```{r} -->
<!-- ttest_results_matrix_for_plot$is_control <- "No" -->
<!-- rs_in_control <- rownames(ttest_results_matrix_for_plot)[gsub(pattern = "-.*", replacement = "", x = rownames(ttest_results_matrix_for_plot)) %in% control_table$rsID.] -->
<!-- ttest_results_matrix_for_plot[rs_in_control, "is_control"] <- "Yes" -->
<!-- ttest_results_matrix_for_plot$name <- rownames(ttest_results_matrix_for_plot) -->
<!-- ttest_results_matrix_for_plot$is_control <- as.factor(ttest_results_matrix_for_plot$is_control) -->
<!-- ttest_results_matrix_for_plot$is_control_and_significant_sample_101  <- as.factor(ifelse((ttest_results_matrix_for_plot$`101_student_p_value` < 0.05 & ttest_results_matrix_for_plot$is_control == "Yes"), TRUE, FALSE)) -->

<!-- ggplot(ttest_results_matrix_for_plot, aes(x = `101_student_estimate_value`, y = -log10(`101_student_p_value`), color = is_control_and_significant_sample_101)) + geom_point() + facet_wrap(~ is_control_and_significant_sample_101) + -->
<!--   ggtitle("Volcano plot of the -log10 of Student test's p-value ") +  -->
<!--   ylab("-log10(adjusted combined p-value of Student test)") + -->
<!--   xlab("Estimated difference between means") +  -->
<!--   labs(color = "Is a control and significant sequence?") + -->
<!--   scale_colour_discrete(labels = paste(levels(ttest_results_matrix_for_plot$is_control_and_significant_sample_101), table(ttest_results_matrix_for_plot$is_control_and_significant_sample_101))) -->
<!-- ``` -->

<!-- The control sequences that have a significant p-values are -->

<!-- ```{r} -->
<!-- rownames(ttest_results_matrix_for_plot[ttest_results_matrix_for_plot$is_control_and_significant_sample_101 == TRUE,]) -->
<!-- ``` -->

<!-- #### Cochran-Mantel-Haenszel chi-squared test for HSO sample -->

<!-- The following test will help to determine if there is an effect of location and allele on the transcription. -->

<!-- ```{r} -->
<!-- chi_two_array <- array(c(sum(ttest_results_matrix_for_plot[grep("REF-Left", rownames(ttest_results_matrix_for_plot)), "101_student_p_value"] < 0.05), -->
<!--                          sum(ttest_results_matrix_for_plot[grep("REF-Left", rownames(ttest_results_matrix_for_plot)), "101_student_p_value"] > 0.05), -->
<!--                          sum(ttest_results_matrix_for_plot[grep("ALT-Left", rownames(ttest_results_matrix_for_plot)), "101_student_p_value"] < 0.05), -->
<!--                          sum(ttest_results_matrix_for_plot[grep("ALT-Left", rownames(ttest_results_matrix_for_plot)), "101_student_p_value"] > 0.05), -->
<!--                          sum(ttest_results_matrix_for_plot[grep("REF-Middle", rownames(ttest_results_matrix_for_plot)), "101_student_p_value"] < 0.05), -->
<!--                          sum(ttest_results_matrix_for_plot[grep("REF-Middle", rownames(ttest_results_matrix_for_plot)), "101_student_p_value"] > 0.05), -->
<!--                          sum(ttest_results_matrix_for_plot[grep("ALT-Middle", rownames(ttest_results_matrix_for_plot)), "101_student_p_value"] < 0.05), -->
<!--                          sum(ttest_results_matrix_for_plot[grep("ALT-Middle", rownames(ttest_results_matrix_for_plot)), "101_student_p_value"] > 0.05), -->
<!--                          sum(ttest_results_matrix_for_plot[grep("REF-Right", rownames(ttest_results_matrix_for_plot)), "101_student_p_value"] < 0.05), -->
<!--                          sum(ttest_results_matrix_for_plot[grep("REF-Right", rownames(ttest_results_matrix_for_plot)), "101_student_p_value"] > 0.05), -->
<!--                          sum(ttest_results_matrix_for_plot[grep("ALT-Right", rownames(ttest_results_matrix_for_plot)), "101_student_p_value"] < 0.05), -->
<!--                          sum(ttest_results_matrix_for_plot[grep("ALT-Right", rownames(ttest_results_matrix_for_plot)), "101_student_p_value"] > 0.05)), -->
<!--                        dim = c(2,2,3), -->
<!--                        dimnames = list(Significant = c("Yes", "No"), -->
<!--                                        Allele = c("REF", "ALT"), -->
<!--                                        Location = c("Left", "Middle", "Right"))) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- chi_two_df <- data.frame(signif = c(sum(ttest_results_matrix_for_plot[grep("ALT-Left", rownames(ttest_results_matrix_for_plot)), "101_student_p_value"] < 0.05), -->
<!--                                     sum(ttest_results_matrix_for_plot[grep("ALT-Middle", rownames(ttest_results_matrix_for_plot)), "101_student_p_value"] < 0.05), -->
<!--                                     sum(ttest_results_matrix_for_plot[grep("ALT-Right", rownames(ttest_results_matrix_for_plot)), "101_student_p_value"] < 0.05)), -->
<!--                          non_signif = c(sum(ttest_results_matrix_for_plot[grep("ALT-Left", rownames(ttest_results_matrix_for_plot)), "101_student_p_value"] > 0.05), -->
<!--                                         sum(ttest_results_matrix_for_plot[grep("ALT-Middle", rownames(ttest_results_matrix_for_plot)), "101_student_p_value"] > 0.05), -->
<!--                                         sum(ttest_results_matrix_for_plot[grep("ALT-Right", rownames(ttest_results_matrix_for_plot)), "101_student_p_value"] > 0.05))) -->
<!-- rownames(chi_two_df) <- c("Left", "Middle", "Right") -->
<!-- chi_two_df$location = rownames(chi_two_df) -->
<!-- chi_two_df_melt <- melt(chi_two_df) -->
<!-- ggplot(chi_two_df_melt, aes(x = location, y = value, fill = location)) + geom_bar(stat = "identity") +  -->
<!--   facet_wrap(~ variable) + -->
<!--   ggtitle("Barplot of the number of significant combined adjusted p-value\nfor sample HSO on ALT sequences") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- chi_two_df <- data.frame(signif = c(sum(ttest_results_matrix_for_plot[grep("REF-Left", rownames(ttest_results_matrix_for_plot)), "101_student_p_value"] < 0.05), -->
<!--                                     sum(ttest_results_matrix_for_plot[grep("REF-Middle", rownames(ttest_results_matrix_for_plot)), "101_student_p_value"] < 0.05), -->
<!--                                     sum(ttest_results_matrix_for_plot[grep("REF-Right", rownames(ttest_results_matrix_for_plot)), "101_student_p_value"] < 0.05)), -->
<!--                          non_signif = c(sum(ttest_results_matrix_for_plot[grep("REF-Left", rownames(ttest_results_matrix_for_plot)), "101_student_p_value"] > 0.05), -->
<!--                                         sum(ttest_results_matrix_for_plot[grep("REF-Middle", rownames(ttest_results_matrix_for_plot)), "101_student_p_value"] > 0.05), -->
<!--                                         sum(ttest_results_matrix_for_plot[grep("REF-Right", rownames(ttest_results_matrix_for_plot)), "101_student_p_value"] > 0.05))) -->
<!-- rownames(chi_two_df) <- c("Left", "Middle", "Right") -->
<!-- chi_two_df$location = rownames(chi_two_df) -->
<!-- chi_two_df_melt <- melt(chi_two_df) -->
<!-- ggplot(chi_two_df_melt, aes(x = location, y = value, fill = location)) + geom_bar(stat = "identity") +  -->
<!--   facet_wrap(~ variable) + -->
<!--   ggtitle("Barplot of the number of significant combined adjusted p-value\nfor sample HSO on REF sequences") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- ftable(chi_two_array) -->

<!-- chi_two_array <- array(c(220, 3727, 227, 3824, -->
<!--                          200, 3747, 207, 3844, -->
<!--                          198, 3749, 206, 3845), -->
<!--                        dim = c(2,2,3), -->
<!--                        dimnames = list(Significant = c("Yes", "No"), -->
<!--                                        Allele = c("REF", "ALT"), -->
<!--                                        Location = c("Left", "Middle", "Right"))) -->
<!-- # mantelhaen.test(chi_two_array) -->
<!-- ``` -->

<!-- We can look at the odds ratio to have a have an idea of the future result -->

<!-- ```{r} -->
<!-- apply(chi_two_array, 3, odds.ratio) -->
<!-- ``` -->

<!-- The odds ratio are very close. We can think that the significance of the Fisher test on REF or ALT sequence is not dependent to the sequence shifting. -->

<!-- And here is the chi-squared results -->

<!-- ```{r} -->
<!-- mantelhaen_results <- mantelhaen.test(chi_two_array) -->
<!-- mantelhaen_results_df <- data.frame(p.value = mantelhaen_results$p.value, odd_ratio = mantelhaen_results$estimate) -->
<!-- rownames(mantelhaen_results_df) <- "" -->
<!-- kable(mantelhaen_results_df) -->
<!-- ``` -->

<!-- The p-value is not significant, as expected above. -->

### Results for HSO sample

<!-- # ```{r} -->
<!-- # ggplot(ttest_results_matrix_for_plot, aes(x = HSO_student_estimate_value, y = -log10(HSO_student_p_value), color = location)) +  -->
<!-- #   geom_point() + facet_wrap(~ location) + -->
<!-- #   ggtitle("Violin plot of the p-values of the Student test and the estimated difference of means,\nsplitted by location for sample HSO") -->
<!-- # ggplot(ttest_results_matrix_for_plot, aes(x = HSO_student_estimate_value, y = -log10(HSO_student_p_value), color = allele)) +  -->
<!-- #   geom_point() + facet_wrap(~ allele) + -->
<!-- #   ggtitle("Violin plot of the p-values of the Student test and the estimated difference of means,\nsplitted by allele for sample HSO") -->
<!-- # ``` -->
<!-- #  -->
<!-- # The violin plots above show the distribution of Student test's p-values among each sequence for sample HSO. I draw those plots to try to see if there is a major difference between location and/or allele. We can't detect a real difference here. -->

```{r}
hist(ttest_results_matrix_for_plot$HSO_student_p_value, main = "Distribution of raw p-value of the Student test", ylab = "Student's test p-values")
hist(ttest_results_matrix_for_plot$HSO_student_adj_p_value, main = "Distribution of adjusted p-value of the Student test", ylab = "Student's test adjusted p-values")
```

When looking at the adjusted p-values (by the Benjamini Hochberg method), none of them are significant (i.e lower than 0.05).

<!-- #### List of significant sequences for the Student test (i.e p-value lower than 0.05) for sample HSO -->

<!-- There are `r sum(ttest_results_matrix_for_plot$HSO_student_p_value < 0.05)` sequences with a significant p-value for the Student test. The list can be obtained using the table "Table_Student_tests_chapter_2.6.csv" or below as an interactive table -->

<!-- ```{r} -->
<!-- datatable(data.frame(sequence_name = rownames(ttest_results_matrix_for_plot[ttest_results_matrix_for_plot$HSO_student_p_value<0.05,]))) -->
<!-- ``` -->

<!-- The following violin plot shows the number of sequences that are control sequences and also have a significant p-value for the Student test. -->

<!-- ```{r} -->
<!-- ttest_results_matrix_for_plot$is_control <- "No" -->
<!-- rs_in_control <- rownames(ttest_results_matrix_for_plot)[gsub(pattern = "-.*", replacement = "", x = rownames(ttest_results_matrix_for_plot)) %in% control_table$rsID.] -->
<!-- ttest_results_matrix_for_plot[rs_in_control, "is_control"] <- "Yes" -->
<!-- ttest_results_matrix_for_plot$name <- rownames(ttest_results_matrix_for_plot) -->
<!-- ttest_results_matrix_for_plot$is_control <- as.factor(ttest_results_matrix_for_plot$is_control) -->
<!-- ttest_results_matrix_for_plot$is_control_and_significant_sample_HSO  <- as.factor(ifelse((ttest_results_matrix_for_plot$HSO_student_p_value < 0.05 & ttest_results_matrix_for_plot$is_control == "Yes"), TRUE, FALSE)) -->

<!-- ggplot(ttest_results_matrix_for_plot, aes(x = HSO_student_estimate_value, y = -log10(HSO_student_p_value), color = is_control_and_significant_sample_HSO)) + geom_point() + facet_wrap(~ is_control_and_significant_sample_HSO) + -->
<!--   ggtitle("Volcano plot of the -log10 of Student test's p-value ") +  -->
<!--   ylab("-log10(adjusted combined p-value of Student test)") + -->
<!--   xlab("Estimated difference between means") +  -->
<!--   labs(color = "Is a control and significant sequence?") + -->
<!--   scale_colour_discrete(labels = paste(levels(ttest_results_matrix_for_plot$is_control_and_significant_sample_HSO), table(ttest_results_matrix_for_plot$is_control_and_significant_sample_HSO))) -->
<!-- ``` -->

<!-- The control sequences that have a significant p-values are -->

<!-- ```{r} -->
<!-- rownames(ttest_results_matrix_for_plot[ttest_results_matrix_for_plot$is_control_and_significant_sample_HSO == TRUE,]) -->
<!-- ``` -->

<!-- #### Cochran-Mantel-Haenszel chi-squared test for HSO sample -->

<!-- The following test will help to determine if there is an effect of location and allele on the transcription. -->

<!-- ```{r} -->
<!-- chi_two_array <- array(c(sum(ttest_results_matrix_for_plot[grep("REF-Left", rownames(ttest_results_matrix_for_plot)), "HSO_student_p_value"] < 0.05), -->
<!--                          sum(ttest_results_matrix_for_plot[grep("REF-Left", rownames(ttest_results_matrix_for_plot)), "HSO_student_p_value"] > 0.05), -->
<!--                          sum(ttest_results_matrix_for_plot[grep("ALT-Left", rownames(ttest_results_matrix_for_plot)), "HSO_student_p_value"] < 0.05), -->
<!--                          sum(ttest_results_matrix_for_plot[grep("ALT-Left", rownames(ttest_results_matrix_for_plot)), "HSO_student_p_value"] > 0.05), -->
<!--                          sum(ttest_results_matrix_for_plot[grep("REF-Middle", rownames(ttest_results_matrix_for_plot)), "HSO_student_p_value"] < 0.05), -->
<!--                          sum(ttest_results_matrix_for_plot[grep("REF-Middle", rownames(ttest_results_matrix_for_plot)), "HSO_student_p_value"] > 0.05), -->
<!--                          sum(ttest_results_matrix_for_plot[grep("ALT-Middle", rownames(ttest_results_matrix_for_plot)), "HSO_student_p_value"] < 0.05), -->
<!--                          sum(ttest_results_matrix_for_plot[grep("ALT-Middle", rownames(ttest_results_matrix_for_plot)), "HSO_student_p_value"] > 0.05), -->
<!--                          sum(ttest_results_matrix_for_plot[grep("REF-Right", rownames(ttest_results_matrix_for_plot)), "HSO_student_p_value"] < 0.05), -->
<!--                          sum(ttest_results_matrix_for_plot[grep("REF-Right", rownames(ttest_results_matrix_for_plot)), "HSO_student_p_value"] > 0.05), -->
<!--                          sum(ttest_results_matrix_for_plot[grep("ALT-Right", rownames(ttest_results_matrix_for_plot)), "HSO_student_p_value"] < 0.05), -->
<!--                          sum(ttest_results_matrix_for_plot[grep("ALT-Right", rownames(ttest_results_matrix_for_plot)), "HSO_student_p_value"] > 0.05)), -->
<!--                        dim = c(2,2,3), -->
<!--                        dimnames = list(Significant = c("Yes", "No"), -->
<!--                                        Allele = c("REF", "ALT"), -->
<!--                                        Location = c("Left", "Middle", "Right"))) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- chi_two_df <- data.frame(signif = c(sum(ttest_results_matrix_for_plot[grep("ALT-Left", rownames(ttest_results_matrix_for_plot)), "HSO_student_p_value"] < 0.05), -->
<!--                                     sum(ttest_results_matrix_for_plot[grep("ALT-Middle", rownames(ttest_results_matrix_for_plot)), "HSO_student_p_value"] < 0.05), -->
<!--                                     sum(ttest_results_matrix_for_plot[grep("ALT-Right", rownames(ttest_results_matrix_for_plot)), "HSO_student_p_value"] < 0.05)), -->
<!--                          non_signif = c(sum(ttest_results_matrix_for_plot[grep("ALT-Left", rownames(ttest_results_matrix_for_plot)), "HSO_student_p_value"] > 0.05), -->
<!--                                         sum(ttest_results_matrix_for_plot[grep("ALT-Middle", rownames(ttest_results_matrix_for_plot)), "HSO_student_p_value"] > 0.05), -->
<!--                                         sum(ttest_results_matrix_for_plot[grep("ALT-Right", rownames(ttest_results_matrix_for_plot)), "HSO_student_p_value"] > 0.05))) -->
<!-- rownames(chi_two_df) <- c("Left", "Middle", "Right") -->
<!-- chi_two_df$location = rownames(chi_two_df) -->
<!-- chi_two_df_melt <- melt(chi_two_df) -->
<!-- ggplot(chi_two_df_melt, aes(x = location, y = value, fill = location)) + geom_bar(stat = "identity") +  -->
<!--   facet_wrap(~ variable) + -->
<!--   ggtitle("Barplot of the number of significant combined adjusted p-value\nfor sample HSO on ALT sequences") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- chi_two_df <- data.frame(signif = c(sum(ttest_results_matrix_for_plot[grep("REF-Left", rownames(ttest_results_matrix_for_plot)), "HSO_student_p_value"] < 0.05), -->
<!--                                     sum(ttest_results_matrix_for_plot[grep("REF-Middle", rownames(ttest_results_matrix_for_plot)), "HSO_student_p_value"] < 0.05), -->
<!--                                     sum(ttest_results_matrix_for_plot[grep("REF-Right", rownames(ttest_results_matrix_for_plot)), "HSO_student_p_value"] < 0.05)), -->
<!--                          non_signif = c(sum(ttest_results_matrix_for_plot[grep("REF-Left", rownames(ttest_results_matrix_for_plot)), "HSO_student_p_value"] > 0.05), -->
<!--                                         sum(ttest_results_matrix_for_plot[grep("REF-Middle", rownames(ttest_results_matrix_for_plot)), "HSO_student_p_value"] > 0.05), -->
<!--                                         sum(ttest_results_matrix_for_plot[grep("REF-Right", rownames(ttest_results_matrix_for_plot)), "HSO_student_p_value"] > 0.05))) -->
<!-- rownames(chi_two_df) <- c("Left", "Middle", "Right") -->
<!-- chi_two_df$location = rownames(chi_two_df) -->
<!-- chi_two_df_melt <- melt(chi_two_df) -->
<!-- ggplot(chi_two_df_melt, aes(x = location, y = value, fill = location)) + geom_bar(stat = "identity") +  -->
<!--   facet_wrap(~ variable) + -->
<!--   ggtitle("Barplot of the number of significant combined adjusted p-value\nfor sample HSO on REF sequences") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- ftable(chi_two_array) -->

<!-- chi_two_array <- array(c(259, 3688, 267, 3784, -->
<!--                          233, 3714, 239, 3812, -->
<!--                          224, 3723, 230, 3821), -->
<!--                        dim = c(2,2,3), -->
<!--                        dimnames = list(Significant = c("Yes", "No"), -->
<!--                                        Allele = c("REF", "ALT"), -->
<!--                                        Location = c("Left", "Middle", "Right"))) -->
<!-- # mantelhaen.test(chi_two_array) -->
<!-- ``` -->

<!-- We can look at the odds ratio to have a have an idea of the future result -->

<!-- ```{r} -->
<!-- apply(chi_two_array, 3, odds.ratio) -->
<!-- ``` -->

<!-- The odds ratio are very close. We can think that the significance of the Fisher test on REF or ALT sequence is not dependent to the sequence shifting. -->

<!-- And here is the chi-squared results -->

<!-- ```{r} -->
<!-- mantelhaen_results <- mantelhaen.test(chi_two_array) -->
<!-- mantelhaen_results_df <- data.frame(p.value = mantelhaen_results$p.value, odd_ratio = mantelhaen_results$estimate) -->
<!-- rownames(mantelhaen_results_df) <- "" -->
<!-- kable(mantelhaen_results_df) -->
<!-- ``` -->

<!-- The p-value is not significant, as expected above. -->

### Export table

For this particular Student test, I can't create a table containing counts.

I created and exported a table containing all results for this pool of Fisher tests. The table contains:

* The estimated difference between means for both samples
* p-value of Student test of comparison for both samples
* Adjusted p-value of Fisher test using the Benjamini-Hochberg correction method for multiple testing for both samples
* Qvalue of adjusted p-value for both samples

```{r}
write.csv(x = ttest_results_matrix_for_plot,
          file = "/data/nin/Workspace/Runs/Run_384/Tables/Table_Student_tests_chapter_2.6.csv", 
          row.names = TRUE)
```

<!-- ### Qvalues -->

<!-- I computed the qvalues of adjusted p-values.The following histogram shows the distribution of adjusted p-value and draw a line for the qvalues and the local FDR.  -->

<!-- ```{r} -->
<!-- # hist(qvalue(ttest_results_matrix_for_plot$student_p_value)) -->

<!-- ggplot(ttest_results_matrix_for_plot, aes(x = student_estimate_value, y = -log10(student_p_value), color = qval_adj_pvalue_FDR_5)) + geom_point() + facet_wrap(~ qval_adj_pvalue_FDR_5) + -->
<!--   ggtitle("Volcano plot of the -log10 of Student test's p-value ") +  -->
<!--   ylab("-log10(adjusted combined p-value of Student test)") + -->
<!--   xlab("Estimated difference between means") +  -->
<!--   labs(color = "Is qvalue significant at FDR 5%?") + -->
<!--   scale_colour_discrete(labels = paste(levels(ttest_results_matrix_for_plot$qval_adj_pvalue_FDR_5), table(ttest_results_matrix_for_plot$qval_adj_pvalue_FDR_5))) -->

<!-- ggplot(ttest_results_matrix_for_plot, aes(x = student_estimate_value, y = -log10(student_p_value), color = qval_adj_pvalue_FDR_10)) + geom_point() + facet_wrap(~ qval_adj_pvalue_FDR_10) + -->
<!--   ggtitle("Volcano plot of the -log10 of Student test's p-value ") +  -->
<!--   ylab("-log10(adjusted combined p-value of Student test)") + -->
<!--   xlab("Estimated difference between means") +  -->
<!--   labs(color = "Is qvalue significant at FDR 10%?") + -->
<!--   scale_colour_discrete(labels = paste(levels(ttest_results_matrix_for_plot$qval_adj_pvalue_FDR_10), table(ttest_results_matrix_for_plot$qval_adj_pvalue_FDR_10))) -->
<!-- ``` -->