---
output: html_document
editor_options: 
  chunk_output_type: console
---

## Fisher test to mesure the influence of ALT allele versus REF allele on cDNA and gDNA expression level, splitted by location (Left, Middle and Right)

On the following analysis, we focused on measuring the effect of the ALT allele versus the REF allele on transcription. This analysis will be done separately for each sequence while taking into account the location (Left, Middle, Right).

This was perform by creating a contingency matrix of the count of the ALT and the REF allele in the cDNA and the gDNA sample. Contingency matrices look like the following:

|     | cDNA | gDNA |
|-----|------|------|
| rs9999446-ALT-Left-1 | 138   | 74 |
| rs9999446-REF-Left-1 | 221   | 63 |

This example is for the comparison of rs9999446-Left-1 and rs9999446-REF-Left-1 101_replicat_1.

```{r}
# Function to create contingency matrix to perform fisher test
create_cont_matrix_and_run_fisher_test_variant_vs_ref <- function(sequence_name, samplename, count_mat){
  sequence_name_ref_tmp <- gsub(pattern = "ALT", replacement = "REF", x = sequence_name)
  # Change the number at the end of the sequence by -1 to get the REF number.
  sequence_name_ref <- gsub(pattern = "-2|-3", replacement = "-1", x = sequence_name_ref_tmp)  
  cont_matrix <- matrix(data = NA, nrow = 2, ncol = 2)
  rownames(cont_matrix) <- c("Alt", "Ref")
  colnames(cont_matrix) <- c("cDNA", "gDNA")
  cont_matrix["Alt", "cDNA"] <- count_mat[sequence_name, samplename]
  cont_matrix["Ref", "cDNA"] <- count_mat[sequence_name_ref, samplename]
  cont_matrix["Alt", "gDNA"] <- count_mat[sequence_name, gsub(pattern = "cDNA", replacement = "ADNg", x = samplename)]
  cont_matrix["Ref", "gDNA"] <- count_mat[sequence_name_ref, gsub(pattern = "cDNA", replacement = "ADNg", x = samplename)]
  # Make on fisher test and extract value
  fisher_test <- fisher.test(x = cont_matrix, alternative = "two.sided")
  ficher_test_pval <- fisher_test$p.value
  ficher_test_oddratio <- fisher_test$estimate
  # fisher_results[sequence_name, gsub(pattern = "_cDNA", replacement = "", 
  # x = samplename)] <- fisher_test$p.value
  # fisher_odd_ratio[sequence_name, gsub(pattern = "_cDNA", replacement = "", 
  # x = samplename)] <- fisher_test$estimate 
  return(list(pval = ficher_test_pval, oddratio = ficher_test_oddratio))
}
```

```{r statistical_tests}
# Prepare the result table
rs_unique_names <- unique(gsub(pattern = "-REF", replacement = "", 
                               x = gsub(pattern = "-ALT", replacement = "", rownames(count_matrix))))

fisher_results <- matrix(data = NA, nrow = length(rs_unique_names), ncol = length(colnames_order_cDNA))
rownames(fisher_results) <- rs_unique_names
colnames(fisher_results) <- gsub(pattern = "cDNA_", replacement = "", x = colnames_order_cDNA)
# fisher_odd_ratio <- matrix(data = NA, nrow = length(rs_unique_names), ncol = length(colnames_order_cDNA))
fisher_odd_ratio <- fisher_results
# Loop over the number of possible variant for the same sequence
for(Alt in 1:3){
  # Get the number of the variant
  grep_ref = paste0("-", Alt)
  # 
  Alt_names_loop <- Alt_names[grep(grep_ref, Alt_names)]
  Ref_names_loop <- gsub(pattern = grep_ref, replacement = "-1",
                         x = gsub(pattern = "ALT", replacement = "REF", x = Alt_names_loop))
  
  Alt_cDNA_matrix_loop <- count_matrix[Alt_names_loop, colnames_order_cDNA]
  Ref_cDNA_matrix_loop <- count_matrix[Ref_names_loop, colnames_order_cDNA]
  Alt_gDNA_matrix_loop <- count_matrix[Alt_names_loop, colnames_order_gDNA]
  Ref_gDNA_matrix_loop <- count_matrix[Ref_names_loop, colnames_order_gDNA]
  
  # Loop to create contingency table and run the fischer test
  for(sample in colnames_order_cDNA){
    for(i in 1:nrow(Alt_cDNA_matrix_loop)){
      rs_name <- rownames(Alt_cDNA_matrix_loop)[i]
      # rs_name <- gsub(pattern = "-ALT", replacement = "", rownames(Alt_cDNA_matrix_loop)[i])
      fisher_test_loop <- create_cont_matrix_and_run_fisher_test_variant_vs_ref(sequence_name = rs_name, samplename = sample, count_mat = count_matrix)
      fisher_results[gsub(pattern = "ALT-", replacement = "", x = rs_name),
                     gsub(pattern = "_cDNA", replacement = "", x = sample)] <-fisher_test_loop$pval
      fisher_odd_ratio[gsub(pattern = "ALT-", replacement = "", x = rs_name),
                       gsub(pattern = "_cDNA", replacement = "", x = sample)] <- fisher_test_loop$oddratio
      # cont_matrix <- matrix(data = NA, nrow = 2, ncol = 2)
      # rownames(cont_matrix) <- c("Alt", "Ref")
      # colnames(cont_matrix) <- c("cDNA", "gDNA")
      # Alt_cDNA <- Alt_cDNA_matrix_loop[i, sample]
      # Ref_cDNA <- Ref_cDNA_matrix_loop[i, sample]
      # Alt_gDNA <- Alt_gDNA_matrix_loop[i, gsub(pattern = "cDNA", replacement = "ADNg", x = sample)]
      # Ref_gDNA <- Ref_gDNA_matrix_loop[i, gsub(pattern = "cDNA", replacement = "ADNg", x = sample)]
      # # Fill the Ref-cDNA case
      # cont_matrix[1,1] <- Alt_cDNA
      # # Fill the Alt-cDNA case
      # cont_matrix[2,1] <- Ref_cDNA
      # # Fill the Ref_gDNA case
      # cont_matrix[1,2] <- Alt_gDNA
      # # Fill the Alt gDNA case
      # cont_matrix[2,2] <- Ref_gDNA
      # # Make on fisher test and extract value
      # fisher_test <- fisher.test(x = t(cont_matrix), alternative = "two.sided")
      # fisher_results[rs_name, gsub(pattern = "_cDNA", replacement = "", 
      #                              x = sample)] <- fisher_test$p.value
      # fisher_odd_ratio[rs_name, gsub(pattern = "_cDNA", replacement = "", 
      #                                x = sample)] <- fisher_test$estimate      
      # hypergeom_results[rs_name, gsub(pattern = "_cDNA", replacement = "", 
      #                                 x = sample)] <- phyper(q = Alt_cDNA - 1,
      #                                                        m = sum(Alt_cDNA, Alt_gDNA),
      #                                                        n = sum(Ref_cDNA, Ref_gDNA),
      #                                                        k = sum(Ref_cDNA, Alt_cDNA),
      #                                                        lower.tail = F)
    }
  }
}

# pairs(fisher_results, lower.panel = dens_plotFun, upper.panel = NULL, main = "Fisher test p-value for each condition")
```

After performing the all the Fisher tests, we can draw the following plot colored by "location" (i.e Left/Middle/Right).

On this plot, the lower triangle contains scatter plot of -log10(p-value) paired by samples. The diagonal represent the distribution of the -log10(p-value) and the upper triangle shows the pearson correlation between -log10(p-value) of each sample (gray) and for each sample splitted by "location" (red/green/blue).

```{r fisher_test_plots_log10_altvsref}
# ```{r fisher_test_plots_altvsref}
# fisher_results_df <- as.data.frame(fisher_results)
# # Replace NA value with a very low value to avoir infinite value when transforming into log10.
# fisher_results_df[is.na(fisher_results_df)] <- 2.737124e-321
# fisher_results_df$location <- gsub(pattern = "rs[0-9]*-|-[123]", replacement = "", x = rownames(fisher_results_df))
# ggpairs(fisher_results_df[1:4], aes(color = fisher_results_df$location)) + ggtitle("Plot matrix for raw fisher test p-value")
# ```

fisher_results_df_log <- as.data.frame(-log10(fisher_results))
# fisher_results_df_log[is.na(fisher_results_df_log)] <- 2.737124e-321
fisher_results_df_log$location <- gsub(pattern = "rs[0-9]*-|-[123]", replacement = "", x = rownames(fisher_results_df_log))
fisher_results_df_log_highly_significant_all_sample <- fisher_results_df_log[(fisher_results_df_log$`101_replicat_1` > 6 & fisher_results_df_log$`101_replicat_2` > 6) | (fisher_results_df_log$HSO_replicat_1 > 6 & fisher_results_df_log$HSO_replicat_2 > 6),]
# fisher_results_df_log[fisher_results_df_log$`101_replicat_1` > 6 & fisher_results_df_log$`101_replicat_2` > 6 & fisher_results_df_log$HSO_replicat_1 > 6 & fisher_results_df_log$HSO_replicat_2 > 6, "highly_significant"] <- rownames(fisher_results_df_log[fisher_results_df_log$`101_replicat_1` > 6 & fisher_results_df_log$`101_replicat_2` > 6 & fisher_results_df_log$HSO_replicat_1 > 6 & fisher_results_df_log$HSO_replicat_2 > 6,])
ggpairs(fisher_results_df_log[,1:4], aes(color = fisher_results_df_log$location)) + ggtitle("Plot matrix for -log10 fisher test p-value")
# ggpairs(fisher_results_df_log[,1:5], aes(color = location), lower = list(continuous = function(data, mapping, ...) ggally_points(data, mapping, ...) + ggrepel::geom_label_repel(aes(label = fisher_results_df_log$highly_significant)))) + ggtitle("Plot matrix for -log10 fisher test p-value")

# ggplot(fisher_results_df_log, aes (x = HSO_replicat_1, y = HSO_replicat_2, color = location)) + geom_point() + geom_vline(xintercept = 20) + geom_hline(yintercept = 20)

# fisher_results_df
# pairs(fisher_results_df_log[,-grep("location", colnames(fisher_results_df_log))], lower.panel = dens_plotFun, upper.panel = NULL, main = "Fisher test p-value for each condition")

# fisher_results_df <- as.data.frame(fisher_results)
# fisher_results_df$location <- gsub(pattern = "rs[0-9]*-|-[123]", replacement = "", x = rownames(fisher_results_df))
# ggplot(fisher_results_df_log, aes (x = HSO_replicat_1, y = HSO_replicat_2, color = location)) + geom_point()
```

### Table of significant p-value

The following table shows the number of p-value lower than 0.01 (i.e log2(p-value) greater than 2). 

```{r}
rep1_101_signif_2 <- sum(fisher_results_df_log$`101_replicat_1` > 2)
rep2_101_signif_2 <- sum(fisher_results_df_log$`101_replicat_2` > 2)
rep1_HSO_signif_2 <- sum(fisher_results_df_log$HSO_replicat_1 > 2)
rep2_HSO_signif_2 <- sum(fisher_results_df_log$HSO_replicat_2 > 2)
signif_df <- as.data.frame(t(data.frame(replicate1_101 = rep1_101_signif_2,
                                        replicate2_101 = rep2_101_signif_2,
                                        replicate1_HSO = rep1_HSO_signif_2,
                                        replicate2_HSO = rep2_HSO_signif_2)))
control_sequence_nb_rep1_101 <- sum(control_table$rsID. %in% gsub(pattern = "-.*", replacement = "", 
                                                                  x = rownames(fisher_results_df_log[fisher_results_df_log$`101_replicat_1` > 2,])))
control_sequence_nb_rep2_101 <- sum(control_table$rsID. %in% gsub(pattern = "-.*", replacement = "", 
                                                                  x = rownames(fisher_results_df_log[fisher_results_df_log$`101_replicat_2` > 2,])))
control_sequence_nb_rep1_HSO <- sum(control_table$rsID. %in% gsub(pattern = "-.*", replacement = "", 
                                                                  x = rownames(fisher_results_df_log[fisher_results_df_log$HSO_replicat_1 > 2,])))
control_sequence_nb_rep2_HSO <- sum(control_table$rsID. %in% gsub(pattern = "-.*", replacement = "", 
                                                                  x = rownames(fisher_results_df_log[fisher_results_df_log$HSO_replicat_2 > 2,])))

colnames(signif_df) <- c("Number_of_p-value_lower_than_0.01")
signif_df$Number_of_significant_control <- c(control_sequence_nb_rep1_101, control_sequence_nb_rep2_101, control_sequence_nb_rep1_HSO, control_sequence_nb_rep2_HSO)
kable(signif_df)

# DT::datatable(t(data.frame(replicate1_101 = rep1_101_signif_2,
#                            replicate2_101 = rep2_101_signif_2,
#                            replicate1_HSO = rep1_HSO_signif_2,
#                            replicate2_HSO = rep2_HSO_signif_2)), colnames = "Number of p-value lower than 0.01")
```

After that, we extracted the most significant p-value with a threshold at 1e-6. On the following plot, only sequence with a Fisher p-value lower than 1e-6 are shown. 

```{r getting_most_significant_p-value_ALT_vs_REF}
fisher_results_df_log_highly_significant_all_sample$is_control <- "No"
rs_in_control <- rownames(fisher_results_df_log_highly_significant_all_sample)[gsub(pattern = "-.*", replacement = "", x = rownames(fisher_results_df_log_highly_significant_all_sample)) %in% control_table$rsID.]
fisher_results_df_log_highly_significant_all_sample[rs_in_control, "is_control"] <- "Yes"
fisher_results_df_log_highly_significant_all_sample$name <- rownames(fisher_results_df_log_highly_significant_all_sample)

# ggpairs(fisher_results_df_log_highly_significant_all_sample[, 1:4], aes(color = as.factor(fisher_results_df_log_highly_significant_all_sample$location)), lower = list(continuous = function(data, mapping, ...) ggally_points(data, mapping, ...) + ggrepel::geom_label_repel(aes(label = fisher_results_df_log_highly_significant_all_sample$name))), upper = NULL) + ggtitle("Plot matrix for -log10 p-value significant in all samples")
fisher_results_df_log_highly_significant_all_sample %>% ggpairs(., legend = 1, columns = 1:4, 
                                                                aes(color = fisher_results_df_log_highly_significant_all_sample$is_control), lower = list(continuous = function(data, mapping, ...) ggally_points(data, mapping, ...) + ggrepel::geom_label_repel(aes(label = fisher_results_df_log_highly_significant_all_sample$name), max.overlaps = 50)), upper = NULL) + 
  theme(legend.position = "bottom") + 
  labs(fill = "is_control")
# Get HSO and 101 sequences with a p-value < 1e-6
# sample_HSO_signif_rs_e6 <- gsub(pattern = "-.*", replacement = "", x = rownames(fisher_results_df_log[fisher_results_df_log$HSO_replicat_1 > 6 & fisher_results_df_log$HSO_replicat_2 > 6,])) 
# sample_101_signif_rs_e6 <- gsub(pattern = "-.*", replacement = "", x = rownames(fisher_results_df_log[fisher_results_df_log$`101_replicat_1` > 6 & fisher_results_df_log$`101_replicat_2` > 6,]))
# 
# sample_HSO_signif_rs_e2 <- gsub(pattern = "-.*", replacement = "", x = rownames(fisher_results_df_log[fisher_results_df_log$HSO_replicat_1 > 2 & fisher_results_df_log$HSO_replicat_2 > 2,])) 
# sample_101_signif_rs_e2 <- gsub(pattern = "-.*", replacement = "", x = rownames(fisher_results_df_log[fisher_results_df_log$`101_replicat_1` > 2 & fisher_results_df_log$`101_replicat_2` > 2,]))

# a <- unique(gsub(pattern = "-.*", replacement = "", x = rownames(fisher_results_df_log)))
# sum(a %in% control_names)

# unique(sample_HSO_signif_rs_e6[sample_HSO_signif_rs_e6 %in% control_table$rsID.])
# unique(sample_101_signif_rs_e6[sample_101_signif_rs_e6 %in% control_table$rsID.])

# control_table[control_table$rsID. %in% sample_HSO_signif_rs_e2[sample_HSO_signif_rs_e2 %in% control_table$rsID.],]
# control_table[control_table$rsID. %in% sample_101_signif_rs_e2[sample_101_signif_rs_e2 %in% control_table$rsID.],]
```

```{r}
fisher_results_samples <- as.data.frame(apply(as.data.frame(fisher_results[,grep("rep", colnames(fisher_results))]), 2,
                                              as.numeric))
# Replace 1 value by 0.9999999 to prevent rowsums to be equal to 2. Sumlog can't perform p-value combination when the sum of p-value is equal to 2
fisher_results_samples[fisher_results_samples == "1"] <- 0.9999999

rownames(fisher_results_samples) <- rownames(fisher_results)
colnames(fisher_results_samples) <- paste0("pvalue_", colnames(fisher_results_samples))
colnames(fisher_odd_ratio) <- paste0("odd_ratio_", colnames(fisher_results))

# Correct p-value using the BH method
# Use the index contained in the results of mt.rawp2adjp to get the location of the sequence in the matrix.
adjusted_pval <- apply(fisher_results_samples, 2, mt.rawp2adjp, proc="BH")
fisher_results_samples$adj_pvalue_101_replicat_1 <- adjusted_pval$pvalue_101_replicat_1$adjp[order(adjusted_pval$pvalue_101_replicat_1$index),][,2]
fisher_results_samples$adj_pvalue_101_replicat_2 <- adjusted_pval$pvalue_101_replicat_2$adjp[order(adjusted_pval$pvalue_101_replicat_2$index),][,2]
fisher_results_samples$adj_pvalue_HSO_replicat_1 <- adjusted_pval$pvalue_HSO_replicat_1$adjp[order(adjusted_pval$pvalue_HSO_replicat_1$index),][,2]
fisher_results_samples$adj_pvalue_HSO_replicat_2 <- adjusted_pval$pvalue_HSO_replicat_2$adjp[order(adjusted_pval$pvalue_HSO_replicat_2$index),][,2]

# Combine fisher test p-value 
# rv_vs_notrs_101_combined_pval <- apply(fisher_results_samples[,grep("adj_pvalue_101", colnames(fisher_results_samples))], 1, sumlog)
# fisher_results_samples <- cbind(fisher_results_samples, combined_adj_pval_101 = do.call(rbind,lapply(rv_vs_notrs_101_combined_pval,function(v){v$p})))
# rv_vs_notrs_HSO_combined_pval <- apply(fisher_results_samples[,grep("adj_pvalue_HSO", colnames(fisher_results_samples))], 1, sumlog)
# fisher_results_samples <- cbind(fisher_results_samples, combined_adj_pval_HSO = do.call(rbind,lapply(rv_vs_notrs_HSO_combined_pval,function(v){v$p})))
rv_vs_notrs_101_combined_pval <- apply(fisher_results_samples[,grep("pvalue_101_replicat", colnames(fisher_results_samples))], 1, sumlog)
fisher_results_samples <- cbind(fisher_results_samples, combined_pval_101 = do.call(rbind,lapply(rv_vs_notrs_101_combined_pval,function(v){v$p})))
rv_vs_notrs_HSO_combined_pval <- apply(fisher_results_samples[,grep("pvalue_HSO_replicat", colnames(fisher_results_samples))], 1, sumlog)
fisher_results_samples <- cbind(fisher_results_samples, combined_pval_HSO = do.call(rbind,lapply(rv_vs_notrs_HSO_combined_pval,function(v){v$p})))

adjusted_pval <- apply(fisher_results_samples[, grep("combined_pval", colnames(fisher_results_samples))], 2, mt.rawp2adjp, proc="BH")
fisher_results_samples$adj_combined_pvalue_101 <- adjusted_pval$combined_pval_101$adjp[order(adjusted_pval$combined_pval_101$index),][,2]
fisher_results_samples$adj_combined_pvalue_HSO <- adjusted_pval$combined_pval_HSO$adjp[order(adjusted_pval$combined_pval_HSO$index),][,2]

# Create the complete table that will be exported
fisher_test_alt_vs_ref_full_table <- cbind(fisher_results_samples, fisher_odd_ratio)
# Check if the odd ratio are concordant between replicates
fisher_test_alt_vs_ref_full_table$concordance_odd_ratio_101 <- FALSE
fisher_test_alt_vs_ref_full_table[(fisher_test_alt_vs_ref_full_table$odd_ratio_101_replicat_1 < 1 & fisher_test_alt_vs_ref_full_table$odd_ratio_101_replicat_2 < 1) |  (fisher_test_alt_vs_ref_full_table$odd_ratio_101_replicat_1 > 1 & fisher_test_alt_vs_ref_full_table$odd_ratio_101_replicat_2 > 1) | (fisher_test_alt_vs_ref_full_table$odd_ratio_101_replicat_1 == 1 & fisher_test_alt_vs_ref_full_table$odd_ratio_101_replicat_2 == 1), "concordance_odd_ratio_101"] <- TRUE
fisher_test_alt_vs_ref_full_table$concordance_odd_ratio_HSO <- FALSE
fisher_test_alt_vs_ref_full_table[(fisher_test_alt_vs_ref_full_table$odd_ratio_HSO_replicat_1 < 1 & fisher_test_alt_vs_ref_full_table$odd_ratio_HSO_replicat_2 < 1) | (fisher_test_alt_vs_ref_full_table$odd_ratio_HSO_replicat_1 > 1 &  fisher_test_alt_vs_ref_full_table$odd_ratio_HSO_replicat_2 > 1) | (fisher_test_alt_vs_ref_full_table$odd_ratio_HSO_replicat_1 == 1 & fisher_test_alt_vs_ref_full_table$odd_ratio_HSO_replicat_2 == 1), "concordance_odd_ratio_HSO"] <- TRUE

fisher_test_alt_vs_ref_full_table$mean_odd_ratio_101 <- log2(rowMeans(fisher_test_alt_vs_ref_full_table[, grep("odd_ratio_101", colnames(fisher_test_alt_vs_ref_full_table))]))
fisher_test_alt_vs_ref_full_table$mean_odd_ratio_HSO <- log2(rowMeans(fisher_test_alt_vs_ref_full_table[, grep("odd_ratio_HSO", colnames(fisher_test_alt_vs_ref_full_table))]))
fisher_test_alt_vs_ref_full_table$concordance_odd_ratio_101 <- as.factor(fisher_test_alt_vs_ref_full_table$concordance_odd_ratio_101)
fisher_test_alt_vs_ref_full_table$concordance_odd_ratio_HSO <- as.factor(fisher_test_alt_vs_ref_full_table$concordance_odd_ratio_HSO)

fisher_test_alt_vs_ref_full_table$qval_adj_combined_pvalue_101 <- qvalue(fisher_test_alt_vs_ref_full_table$adj_combined_pvalue_101)$qvalues
fisher_test_alt_vs_ref_full_table$qval_adj_combined_pvalue_101_FDR_5 <- as.factor(qvalue(fisher_test_alt_vs_ref_full_table$adj_combined_pvalue_101, fdr.level = 0.05)$significant)
fisher_test_alt_vs_ref_full_table$qval_adj_combined_pvalue_101_FDR_10 <- as.factor(qvalue(fisher_test_alt_vs_ref_full_table$adj_combined_pvalue_101, fdr.level = 0.10)$significant)
fisher_test_alt_vs_ref_full_table$qval_adj_combined_pvalue_HSO <- qvalue(fisher_test_alt_vs_ref_full_table$adj_combined_pvalue_HSO)$qvalues
fisher_test_alt_vs_ref_full_table$qval_adj_combined_pvalue_HSO_FDR_5 <- as.factor(qvalue(fisher_test_alt_vs_ref_full_table$adj_combined_pvalue_HSO, fdr.level = 0.05)$significant)
fisher_test_alt_vs_ref_full_table$qval_adj_combined_pvalue_HSO_FDR_10 <- as.factor(qvalue(fisher_test_alt_vs_ref_full_table$adj_combined_pvalue_HSO, fdr.level = 0.10)$significant)
```

### Volcano plot

The following volcano plots show the relation between the p-value and the odd ratio. The plots are splitted depending on the concordance of odd ratio. TRUE indicates concordants odd ratio, whereas FALSE indicante non concordant ones.

```{r volcano_qval_oddratio_alt_vs_ref}
ggplot(fisher_test_alt_vs_ref_full_table, aes(x = mean_odd_ratio_101, y = -log10(adj_combined_pvalue_101), color = concordance_odd_ratio_101)) + 
  geom_point() + 
  facet_wrap(~ concordance_odd_ratio_101) +
  ggtitle("Volcano plot of the -log10 of adjusted combined p-value for samples 101") + 
  ylab("-log10(adjusted combined p-value of Fisher test)") +
  xlab("Mean odd ratio of Fisher test") + 
  labs(color = "Are odd ratio concordant") +
  scale_colour_discrete(labels = paste(levels(fisher_test_alt_vs_ref_full_table$concordance_odd_ratio_101), table(fisher_test_alt_vs_ref_full_table$concordance_odd_ratio_101)))

ggplot(fisher_test_alt_vs_ref_full_table, aes(x = mean_odd_ratio_HSO, y = -log10(adj_combined_pvalue_HSO), color = concordance_odd_ratio_HSO)) + 
  geom_point() + 
  facet_wrap(~ concordance_odd_ratio_HSO) +
  ggtitle("Volcano plot f the -log10 of adjusted combined p-value for samples HSO") + 
  ylab("-log10(adjusted combined p-value of Fisher test)") +
  xlab("Mean odd ratio of Fisher test") + 
  labs(color = "Are odd ratio concordant") +
  scale_colour_discrete(labels = paste(levels(fisher_test_alt_vs_ref_full_table$concordance_odd_ratio_HSO), table(fisher_test_alt_vs_ref_full_table$concordance_odd_ratio_HSO)))
```

### Export table

For this particular Fisher test, I can't create a table containing counts.

I created and exported a table containing all results for this pool of Fisher tests. The table contains:

* Raw counts of each sequence for each samples
* p-value of Fisher test of each sequence for each sample
* Combined p-value using the Fisher method
* Adjusted p-value of Fisher test using the Benjamini-Hochberg correction method for multiple testing
* Combined adjusted p-value using the Fisher method
* The odd ratio of Fisher test for each sequence and each sample
* A boolean column (TRUE of FALSE) for each sample indicating if odd ratio for both replicate are concordant (both < 1 or both = 1 or both > 1)
* The mean odd ratio of each sample
* Qvalue of combined adjusted p-value
* A boolean column identifying significant sequences with a FDR of 5% or 10% (TRUE or FALSE).

```{r}
write.csv(x = fisher_test_alt_vs_ref_full_table,
          file = "/data/nin/Workspace/Runs/Run_384/Tables/Table_Fisher_tests_chapter_2.4.csv", 
          row.names = TRUE)
```

### Table with some sequences having a significant regulatory effect

The following table shows sequences having a significant regulatory effect. The shown sequences are selected by having a combined p-value (using the Fisher combination method for p-value) lower than 0.05

#### Sample 101

```{r}
sample_101_significant_seq_combined_pval <- as.data.frame( rownames(fisher_test_alt_vs_ref_full_table[fisher_test_alt_vs_ref_full_table$adj_combined_pvalue_101 < 0.05,]))
colnames(sample_101_significant_seq_combined_pval) <- "significant_sequences"
datatable(sample_101_significant_seq_combined_pval)

```

#### Sample HSO

```{r}
sample_HSO_significant_seq_combined_pval <- as.data.frame( rownames(fisher_test_alt_vs_ref_full_table[fisher_test_alt_vs_ref_full_table$adj_combined_pvalue_HSO < 0.05,]))
colnames(sample_HSO_significant_seq_combined_pval) <- "significant_sequences"
datatable(sample_HSO_significant_seq_combined_pval)

```

#### Common significant sequences

The following venn diagram shows the number of significant sequences detected in both samples.

```{r}
# Fonction d'aide pour afficher le diagramme de Venn
display_venn <- function(x, ...){
  library(VennDiagram)
  grid.newpage()
  venn_object <- venn.diagram(x, filename = NULL, ...)
  grid.draw(venn_object)
}
display_venn(
  x = c(sample_HSO_significant_seq_combined_pval, sample_101_significant_seq_combined_pval),
  category.names = c(paste0("HSO\nn = ", nrow(sample_HSO_significant_seq_combined_pval)), 
                     paste0("101\nn = ", nrow(sample_101_significant_seq_combined_pval))),
  # Cercles
  lwd = 2,
  lty = 'blank',
  fill = c("#999999", "#E69F00"),
  # Nombres
  cex = .9,
  fontface = "italic",
  # Noms des groupes
  cat.cex = 1,
  cat.fontface = "bold",
  cat.default.pos = "outer",
  cat.dist = c(0.055, 0.055)
)

```

Here is the list of common detected sequences

```{r}
common_significant_seq_combined_pval <- as.data.frame(intersect(sample_101_significant_seq_combined_pval$significant_sequences, sample_HSO_significant_seq_combined_pval$significant_sequences))
colnames(common_significant_seq_combined_pval) <- "Common_significant_sequences"
datatable(common_significant_seq_combined_pval)
```

### Qvalues

I computed the qvalues of adjusted combined p-values for each samples. I draw the following histogram for each sample. I shows the distribution of adjusted combined p-value and draw a line for the qvalues and the local FDR. I also draw 2 volcano plots showing the relation between the p-value and the odd ratio. The plots are splitted depending on the significance of the qvalues. TRUE indicates a significant qvalue, whereas FALSE indicates non significant ones. I draw the plots using a FDR of 5% and 10%. 

#### Sample 101

```{r}
# hist(qvalue(fisher_test_alt_vs_ref_full_table$adj_combined_pvalue_101))

ggplot(fisher_test_alt_vs_ref_full_table, aes(x = mean_odd_ratio_101, y = -log10(adj_combined_pvalue_101), color = qval_adj_combined_pvalue_101_FDR_5)) + 
  geom_point() + 
  facet_wrap(~ qval_adj_combined_pvalue_101_FDR_5) +
  ggtitle("Volcano plot of the -log10 of adjusted combined p-value for samples 101") + 
  ylab("-log10(adjusted combined p-value of Fisher test)") +
  xlab("Mean odd ratio of Fisher test") + 
  labs(color = "Is qvalue significant at FDR 5%?") +
  scale_colour_discrete(labels = paste(levels(fisher_test_alt_vs_ref_full_table$qval_adj_combined_pvalue_101_FDR_5), table(fisher_test_alt_vs_ref_full_table$qval_adj_combined_pvalue_101_FDR_5)))

ggplot(fisher_test_alt_vs_ref_full_table, aes(x = mean_odd_ratio_101, y = -log10(adj_combined_pvalue_101), color = qval_adj_combined_pvalue_101_FDR_10)) + 
  geom_point() + 
  facet_wrap(~ qval_adj_combined_pvalue_101_FDR_10) +
  ggtitle("Volcano plot of the -log10 of adjusted combined p-value for samples 101") + 
  ylab("-log10(adjusted combined p-value of Fisher test)") +
  xlab("Mean odd ratio of Fisher test") + 
  labs(color = "Is qvalue significant at FDR 10%?") +
  scale_colour_discrete(labels = paste(levels(fisher_test_alt_vs_ref_full_table$qval_adj_combined_pvalue_101_FDR_10), table(fisher_test_alt_vs_ref_full_table$qval_adj_combined_pvalue_101_FDR_10)))
```

##### List of significant sequence with a 5% FDR

```{r}
datatable(fisher_test_alt_vs_ref_full_table[fisher_test_alt_vs_ref_full_table$qval_adj_combined_pvalue_101_FDR_5 == TRUE, 
                                            c("adj_combined_pvalue_101", "qval_adj_combined_pvalue_101")])
```

##### List of significant sequence with a 10% FDR

```{r}
datatable(fisher_test_alt_vs_ref_full_table[fisher_test_alt_vs_ref_full_table$qval_adj_combined_pvalue_101_FDR_10 == TRUE, 
                                            c("adj_combined_pvalue_101", "qval_adj_combined_pvalue_101")])
```

#### Sample HSO

```{r}
# hist(qvalue(fisher_test_alt_vs_ref_full_table$adj_combined_pvalue_HSO))

ggplot(fisher_test_alt_vs_ref_full_table, aes(x = mean_odd_ratio_HSO, y = -log10(adj_combined_pvalue_HSO), color = qval_adj_combined_pvalue_HSO_FDR_5)) + 
  geom_point() + 
  facet_wrap(~ qval_adj_combined_pvalue_HSO_FDR_5) +
  ggtitle("Volcano plot of the -log10 of adjusted combined p-value for samples HSO") + 
  ylab("-log10(adjusted combined p-value of Fisher test)") +
  xlab("Mean odd ratio of Fisher test") + 
  labs(color = "Is qvalue significant at FDR 5%?") +
  scale_colour_discrete(labels = paste(levels(fisher_test_alt_vs_ref_full_table$qval_adj_combined_pvalue_HSO_FDR_5), table(fisher_test_alt_vs_ref_full_table$qval_adj_combined_pvalue_HSO_FDR_5)))

ggplot(fisher_test_alt_vs_ref_full_table, aes(x = mean_odd_ratio_HSO, y = -log10(adj_combined_pvalue_HSO), color = qval_adj_combined_pvalue_HSO_FDR_10)) + 
  geom_point() + 
  facet_wrap(~ qval_adj_combined_pvalue_HSO_FDR_10) +
  ggtitle("Volcano plot of the -log10 of adjusted combined p-value for samples HSO") + 
  ylab("-log10(adjusted combined p-value of Fisher test)") +
  xlab("Mean odd ratio of Fisher test") + 
  labs(color = "Is qvalue significant at FDR 10%?") +
  scale_colour_discrete(labels = paste(levels(fisher_test_alt_vs_ref_full_table$qval_adj_combined_pvalue_HSO_FDR_10), table(fisher_test_alt_vs_ref_full_table$qval_adj_combined_pvalue_HSO_FDR_10)))
```

##### List of significant sequence with a 5% FDR

```{r}
datatable(fisher_test_alt_vs_ref_full_table[fisher_test_alt_vs_ref_full_table$qval_adj_combined_pvalue_HSO_FDR_5 == TRUE, 
                                            c("adj_combined_pvalue_HSO", "qval_adj_combined_pvalue_HSO")])
```

##### List of significant sequence with a 10% FDR

```{r}
datatable(fisher_test_alt_vs_ref_full_table[fisher_test_alt_vs_ref_full_table$qval_adj_combined_pvalue_HSO_FDR_10 == TRUE, 
                                            c("adj_combined_pvalue_HSO", "qval_adj_combined_pvalue_HSO")])
```

#### Venn diagrams with qvalues

##### FDR 5%

```{r}
sample_101_significant_seq_qval <- as.data.frame( rownames(fisher_test_alt_vs_ref_full_table[fisher_test_alt_vs_ref_full_table$qval_adj_combined_pvalue_101_FDR_5 == TRUE, 
                                            c("adj_combined_pvalue_101", "qval_adj_combined_pvalue_101")]))
colnames(sample_101_significant_seq_qval) <- "significant_sequences_with_FDR_5"

sample_HSO_significant_seq_qval <- as.data.frame( rownames(fisher_test_alt_vs_ref_full_table[fisher_test_alt_vs_ref_full_table$qval_adj_combined_pvalue_HSO_FDR_5 == TRUE, 
                                            c("adj_combined_pvalue_HSO", "qval_adj_combined_pvalue_HSO")]))
colnames(sample_HSO_significant_seq_qval) <- "significant_sequences_with_FDR_5"

display_venn(
  x = c(sample_HSO_significant_seq_qval, sample_101_significant_seq_qval),
  category.names = c(paste0("HSO\nn = ", nrow(sample_HSO_significant_seq_qval)), 
                     paste0("101\nn = ", nrow(sample_101_significant_seq_qval))),
  # Cercles
  lwd = 2,
  lty = 'blank',
  fill = c("#999999", "#E69F00"),
  # Nombres
  cex = .9,
  fontface = "italic",
  # Noms des groupes
  cat.cex = 1,
  cat.fontface = "bold",
  cat.default.pos = "outer",
  cat.dist = c(0.055, 0.055)
)
```

Here is the list of common detected sequences

```{r}
common_significant_seq_combined_pval <- as.data.frame(intersect(sample_HSO_significant_seq_qval$significant_sequences_with_FDR_5, sample_101_significant_seq_qval$significant_sequences_with_FDR_5))
colnames(common_significant_seq_combined_pval) <- "Common_significant_sequences_with_FDR_5%"
datatable(common_significant_seq_combined_pval)
```

##### FDR 10%

```{r}
sample_101_significant_seq_qval <- as.data.frame( rownames(fisher_test_alt_vs_ref_full_table[fisher_test_alt_vs_ref_full_table$qval_adj_combined_pvalue_101_FDR_10 == TRUE, 
                                            c("adj_combined_pvalue_101", "qval_adj_combined_pvalue_101")]))
colnames(sample_101_significant_seq_qval) <- "significant_sequences_with_FDR_10"

sample_HSO_significant_seq_qval <- as.data.frame( rownames(fisher_test_alt_vs_ref_full_table[fisher_test_alt_vs_ref_full_table$qval_adj_combined_pvalue_HSO_FDR_10 == TRUE, 
                                            c("adj_combined_pvalue_HSO", "qval_adj_combined_pvalue_HSO")]))
colnames(sample_HSO_significant_seq_qval) <- "significant_sequences_with_FDR_10"

display_venn(
  x = c(sample_HSO_significant_seq_qval, sample_101_significant_seq_qval),
  category.names = c(paste0("HSO\nn = ", nrow(sample_HSO_significant_seq_qval)), 
                     paste0("101\nn = ", nrow(sample_101_significant_seq_qval))),
  # Cercles
  lwd = 2,
  lty = 'blank',
  fill = c("#999999", "#E69F00"),
  # Nombres
  cex = .9,
  fontface = "italic",
  # Noms des groupes
  cat.cex = 1,
  cat.fontface = "bold",
  cat.default.pos = "outer",
  cat.dist = c(0.055, 0.055)
)
```

Here is the list of common detected sequences

```{r}
common_significant_seq_combined_pval <- as.data.frame(intersect(sample_HSO_significant_seq_qval$significant_sequences_with_FDR_10, sample_101_significant_seq_qval$significant_sequences_with_FDR_10))
colnames(common_significant_seq_combined_pval) <- "Common_significant_sequences_with_FDR_10%"
datatable(common_significant_seq_combined_pval)
```

<!-- ### Analysis of the impact of sequence shifting -->

<!-- The following analysis aims to detect an impact of the sequence shifting on the transcription.  -->

<!-- For both samples, (101 and HSO), I looked at the adjusted combined p-values of Fisher test obtained when looking at the impact on transcription for each sequences. Before going into statistical tests, I draw some boxplot of p-value on each samples and separated by significance of p-value (i.e adjusted combined p-values lower than 0.05).  -->

<!-- #### Analysis on REF sequences for 101 sample -->

<!-- Here is the boxplot of adjusted combined p-values for sample 101 on REF sequences. -->

<!-- ```{r} -->
<!-- # left_signif <- fisher_test_alt_vs_ref_full_table[grep("REF-Left", rownames(fisher_test_alt_vs_ref_full_table[fisher_test_alt_vs_ref_full_table$qval_adj_combined_pvalue_101 < 0.05,])),] -->

<!-- ref_left_signif <- sum(fisher_test_alt_vs_ref_full_table[grep("Left", rownames(fisher_test_alt_vs_ref_full_table)), "adj_combined_pvalue_101"] < 0.05) -->
<!-- ref_left_nosignif <- sum(fisher_test_alt_vs_ref_full_table[grep("Left", rownames(fisher_test_alt_vs_ref_full_table)), "adj_combined_pvalue_101"] > 0.05) -->
<!-- alt_left_signif <- sum(fisher_test_alt_vs_ref_full_table[grep("Left", rownames(fisher_test_alt_vs_ref_full_table)), "adj_combined_pvalue_101"] < 0.05) -->
<!-- alt_left_nosignif <- sum(fisher_test_alt_vs_ref_full_table[grep("Left", rownames(fisher_test_alt_vs_ref_full_table)), "adj_combined_pvalue_101"] > 0.05) -->
<!-- ref_middle_signif <- sum(fisher_test_alt_vs_ref_full_table[grep("Middle", rownames(fisher_test_alt_vs_ref_full_table)), "adj_combined_pvalue_101"] < 0.05) -->
<!-- ref_middle_nosignif <- sum(fisher_test_alt_vs_ref_full_table[grep("Middle", rownames(fisher_test_alt_vs_ref_full_table)), "adj_combined_pvalue_101"] > 0.05) -->
<!-- alt_middle_signif <- sum(fisher_test_alt_vs_ref_full_table[grep("Middle", rownames(fisher_test_alt_vs_ref_full_table)), "adj_combined_pvalue_101"] < 0.05) -->
<!-- alt_middle_nosignif <- sum(fisher_test_alt_vs_ref_full_table[grep("Middle", rownames(fisher_test_alt_vs_ref_full_table)), "adj_combined_pvalue_101"] > 0.05) -->
<!-- ref_right_signif <- sum(fisher_test_alt_vs_ref_full_table[grep("Right", rownames(fisher_test_alt_vs_ref_full_table)), "adj_combined_pvalue_101"] < 0.05) -->
<!-- ref_right_nosignif <- sum(fisher_test_alt_vs_ref_full_table[grep("Right", rownames(fisher_test_alt_vs_ref_full_table)), "adj_combined_pvalue_101"] > 0.05) -->
<!-- alt_right_signif <- sum(fisher_test_alt_vs_ref_full_table[grep("Right", rownames(fisher_test_alt_vs_ref_full_table)), "adj_combined_pvalue_101"] < 0.05) -->
<!-- alt_right_nosignif <- sum(fisher_test_alt_vs_ref_full_table[grep("Right", rownames(fisher_test_alt_vs_ref_full_table)), "adj_combined_pvalue_101"] > 0.05) -->

<!-- chi_two_array <- array(c(ref_left_signif, ref_left_nosignif, alt_left_signif, alt_left_nosignif, -->
<!--                          ref_middle_signif, ref_middle_nosignif, alt_middle_signif, alt_middle_nosignif, -->
<!--                          ref_right_signif, ref_right_nosignif, alt_right_signif, alt_right_nosignif), -->
<!--                        dim = c(2,2,3), -->
<!--                        dimnames = list(Significant = c("Yes", "No"), -->
<!--                                        Allele = c("REF", "ALT"), -->
<!--                                        Location = c("Left", "Middle", "Right"))) -->
<!-- # marginal_table <- margin.table(chi_two_array, c(2,3)) -->
<!-- # chi_two_array <- array(c(sum(fisher_test_alt_vs_ref_full_table[grep("Left", rownames(fisher_test_alt_vs_ref_full_table)), "adj_combined_pvalue_101"] < 0.05), -->
<!-- #                          sum(fisher_test_alt_vs_ref_full_table[grep("Left", rownames(fisher_test_alt_vs_ref_full_table)), "adj_combined_pvalue_101"] > 0.05), -->
<!-- #                          sum(fisher_test_alt_vs_ref_full_table[grep("Middle", rownames(fisher_test_alt_vs_ref_full_table)), "adj_combined_pvalue_101"] < 0.05), -->
<!-- #                          sum(fisher_test_alt_vs_ref_full_table[grep("Middle", rownames(fisher_test_alt_vs_ref_full_table)), "adj_combined_pvalue_101"] > 0.05), -->
<!-- #                          sum(fisher_test_alt_vs_ref_full_table[grep("Right", rownames(fisher_test_alt_vs_ref_full_table)), "adj_combined_pvalue_101"] < 0.05), -->
<!-- #                          sum(fisher_test_alt_vs_ref_full_table[grep("Right", rownames(fisher_test_alt_vs_ref_full_table)), "adj_combined_pvalue_101"] > 0.05)), -->
<!-- #                        dim = c(2,3), -->
<!-- #                        dimnames = list(Significant = c("Yes", "No"), -->
<!-- #                                        Location = c("Left", "Middle", "Right"))) -->

<!-- # apply(chi_two_array, 3, odds.ratio) -->
<!-- # mantelhaen.test(chi_two_array) -->
<!-- # ftable(chi_two_array) -->
<!-- #  -->
<!-- # chi_two_array <- array(c(2481, 1466, 2483, 1568, -->
<!-- #                          2026, 1921, 2069, 1982, -->
<!-- #                          1761, 2186, 1794, 2257), -->
<!-- #                        dim = c(2,2,3), -->
<!-- #                        dimnames = list(Significant = c("Yes", "No"), -->
<!-- #                                        Allele = c("REF", "ALT"), -->
<!-- #                                        Location = c("Left", "Middle", "Right"))) -->
<!-- # mantelhaen.test(chi_two_array) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- # chi_two_array <- array(c(sum(fisher_test_alt_vs_ref_full_table[grep("Left", rownames(fisher_test_alt_vs_ref_full_table)), "adj_combined_pvalue_101"] < 0.05), -->
<!-- #                          sum(fisher_test_alt_vs_ref_full_table[grep("Middle", rownames(fisher_test_alt_vs_ref_full_table)), "adj_combined_pvalue_101"] < 0.05), -->
<!-- #                          sum(fisher_test_alt_vs_ref_full_table[grep("Right", rownames(fisher_test_alt_vs_ref_full_table)), "adj_combined_pvalue_101"] < 0.05), -->
<!-- #                          sum(fisher_test_alt_vs_ref_full_table[grep("Left", rownames(fisher_test_alt_vs_ref_full_table)), "adj_combined_pvalue_101"] > 0.05), -->
<!-- #                          sum(fisher_test_alt_vs_ref_full_table[grep("Middle", rownames(fisher_test_alt_vs_ref_full_table)), "adj_combined_pvalue_101"] > 0.05), -->
<!-- #                          sum(fisher_test_alt_vs_ref_full_table[grep("Right", rownames(fisher_test_alt_vs_ref_full_table)), "adj_combined_pvalue_101"] > 0.05)), -->
<!-- #                        dim = c(2,3), -->
<!-- #                        dimnames = list(Significant = c("Yes", "No"), -->
<!-- #                                        Location = c("Left", "Middle", "Right"))) -->
<!-- #  -->
<!-- chi_two_df <- data.frame(signif = c(sum(fisher_test_alt_vs_ref_full_table[grep("Left", rownames(fisher_test_alt_vs_ref_full_table)), "adj_combined_pvalue_101"] < 0.05), -->
<!--                                     sum(fisher_test_alt_vs_ref_full_table[grep("Middle", rownames(fisher_test_alt_vs_ref_full_table)), "adj_combined_pvalue_101"] < 0.05), -->
<!--                                     sum(fisher_test_alt_vs_ref_full_table[grep("Right", rownames(fisher_test_alt_vs_ref_full_table)), "adj_combined_pvalue_101"] < 0.05)), -->
<!--                          non_signif = c(sum(fisher_test_alt_vs_ref_full_table[grep("Left", rownames(fisher_test_alt_vs_ref_full_table)), "adj_combined_pvalue_101"] > 0.05), -->
<!--                                         sum(fisher_test_alt_vs_ref_full_table[grep("Middle", rownames(fisher_test_alt_vs_ref_full_table)), "adj_combined_pvalue_101"] > 0.05), -->
<!--                                         sum(fisher_test_alt_vs_ref_full_table[grep("Right", rownames(fisher_test_alt_vs_ref_full_table)), "adj_combined_pvalue_101"] > 0.05))) -->
<!-- rownames(chi_two_df) <- c("Left", "Middle", "Right") -->
<!-- chi_two_df$location = rownames(chi_two_df) -->
<!-- chi_two_df_melt <- melt(chi_two_df) -->
<!-- ggplot(chi_two_df_melt, aes(x = location, y = value, fill = location)) + geom_bar(stat = "identity") +  -->
<!--   facet_wrap(~ variable) + -->
<!--   ggtitle("Barplot of the number of significant adjusted combined p-value\nfor sample 101 on REF sequences")  -->
<!-- # mantelhaen.test(chi_two_df) -->
<!-- ``` -->

<!-- #### Analysis on ALT sequences for 101 sample -->

<!-- Here is the boxplot of adjusted combined p-values for sample 101 on ALT sequences. -->

<!-- ```{r} -->
<!-- chi_two_df <- data.frame(signif = c(sum(fisher_test_alt_vs_ref_full_table[grep("Left", rownames(fisher_test_alt_vs_ref_full_table)), "adj_combined_pvalue_101"] < 0.05), -->
<!--                                     sum(fisher_test_alt_vs_ref_full_table[grep("Middle", rownames(fisher_test_alt_vs_ref_full_table)), "adj_combined_pvalue_101"] < 0.05), -->
<!--                                     sum(fisher_test_alt_vs_ref_full_table[grep("Right", rownames(fisher_test_alt_vs_ref_full_table)), "adj_combined_pvalue_101"] < 0.05)), -->
<!--                          non_signif = c(sum(fisher_test_alt_vs_ref_full_table[grep("Left", rownames(fisher_test_alt_vs_ref_full_table)), "adj_combined_pvalue_101"] > 0.05), -->
<!--                                         sum(fisher_test_alt_vs_ref_full_table[grep("Middle", rownames(fisher_test_alt_vs_ref_full_table)), "adj_combined_pvalue_101"] > 0.05), -->
<!--                                         sum(fisher_test_alt_vs_ref_full_table[grep("Right", rownames(fisher_test_alt_vs_ref_full_table)), "adj_combined_pvalue_101"] > 0.05))) -->
<!-- rownames(chi_two_df) <- c("Left", "Middle", "Right") -->
<!-- chi_two_df$location = rownames(chi_two_df) -->
<!-- chi_two_df_melt <- melt(chi_two_df) -->
<!-- ggplot(chi_two_df_melt, aes(x = location, y = value, fill = location)) + geom_bar(stat = "identity") +  -->
<!--   facet_wrap(~ variable) + -->
<!--   ggtitle("Barplot of the number of significant adjusted combined p-value\nfor sample 101 on ALT sequences") -->
<!-- ``` -->

<!-- #### Cochran-Mantel-Haenszel chi-squared test for sample 101 -->

<!-- We can have the feeling that Left shifting has more significant p-values. We can validate this by running a Cochran-Mantel-Haenszel chi-squared test. I run it on the following table. -->

<!-- ```{r} -->
<!-- ftable(chi_two_array) -->

<!-- chi_two_array <- array(c(2481, 1466, 2483, 1568, -->
<!--                          2026, 1921, 2069, 1982, -->
<!--                          1761, 2186, 1794, 2257), -->
<!--                        dim = c(2,2,3), -->
<!--                        dimnames = list(Significant = c("Yes", "No"), -->
<!--                                        Allele = c("REF", "ALT"), -->
<!--                                        Location = c("Left", "Middle", "Right"))) -->
<!-- ``` -->

<!-- We can look at the odds ratio to have a have an idea of the future result -->

<!-- ```{r} -->
<!-- apply(chi_two_array, 3, odds.ratio) -->
<!-- ``` -->

<!-- The odds ratio are very close. We can think that the significance of the Fisher test on REF or ALT sequence is not dependent to the sequence shifting. -->

<!-- And here is the chi-squared results -->

<!-- ```{r} -->
<!-- mantelhaen_results <- mantelhaen.test(chi_two_array) -->
<!-- mantelhaen_results_df <- data.frame(p.value = mantelhaen_results$p.value, odd_ratio = mantelhaen_results$estimate) -->
<!-- rownames(mantelhaen_results_df) <- "" -->
<!-- kable(mantelhaen_results_df) -->
<!-- ``` -->

<!-- The p-value is not significant, as expected above. -->

<!-- #### Analysis on REF sequences for HSO sample -->

<!-- Here is the boxplot of adjusted combined p-values for sample HSO on REF sequences. -->

<!-- ```{r} -->
<!-- chi_two_array <- array(c(sum(fisher_test_alt_vs_ref_full_table[grep("Left", rownames(fisher_test_alt_vs_ref_full_table)), "adj_combined_pvalue_HSO"] < 0.05), -->
<!--                          sum(fisher_test_alt_vs_ref_full_table[grep("Left", rownames(fisher_test_alt_vs_ref_full_table)), "adj_combined_pvalue_HSO"] > 0.05), -->
<!--                          sum(fisher_test_alt_vs_ref_full_table[grep("Left", rownames(fisher_test_alt_vs_ref_full_table)), "adj_combined_pvalue_HSO"] < 0.05), -->
<!--                          sum(fisher_test_alt_vs_ref_full_table[grep("Left", rownames(fisher_test_alt_vs_ref_full_table)), "adj_combined_pvalue_HSO"] > 0.05), -->
<!--                          sum(fisher_test_alt_vs_ref_full_table[grep("Middle", rownames(fisher_test_alt_vs_ref_full_table)), "adj_combined_pvalue_HSO"] < 0.05), -->
<!--                          sum(fisher_test_alt_vs_ref_full_table[grep("Middle", rownames(fisher_test_alt_vs_ref_full_table)), "adj_combined_pvalue_HSO"] > 0.05), -->
<!--                          sum(fisher_test_alt_vs_ref_full_table[grep("Middle", rownames(fisher_test_alt_vs_ref_full_table)), "adj_combined_pvalue_HSO"] < 0.05), -->
<!--                          sum(fisher_test_alt_vs_ref_full_table[grep("Middle", rownames(fisher_test_alt_vs_ref_full_table)), "adj_combined_pvalue_HSO"] > 0.05), -->
<!--                          sum(fisher_test_alt_vs_ref_full_table[grep("Right", rownames(fisher_test_alt_vs_ref_full_table)), "adj_combined_pvalue_HSO"] < 0.05), -->
<!--                          sum(fisher_test_alt_vs_ref_full_table[grep("Right", rownames(fisher_test_alt_vs_ref_full_table)), "adj_combined_pvalue_HSO"] > 0.05), -->
<!--                          sum(fisher_test_alt_vs_ref_full_table[grep("Right", rownames(fisher_test_alt_vs_ref_full_table)), "adj_combined_pvalue_HSO"] < 0.05), -->
<!--                          sum(fisher_test_alt_vs_ref_full_table[grep("Right", rownames(fisher_test_alt_vs_ref_full_table)), "adj_combined_pvalue_HSO"] > 0.05)), -->
<!--                        dim = c(2,2,3), -->
<!--                        dimnames = list(Significant = c("Yes", "No"), -->
<!--                                        Allele = c("REF", "ALT"), -->
<!--                                        Location = c("Left", "Middle", "Right"))) -->
<!-- # mantelhaen.test(chi_two_array) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- # mantelhaen.test(chi_two_array) -->
<!-- chi_two_df <- data.frame(signif = c(sum(fisher_test_alt_vs_ref_full_table[grep("Left", rownames(fisher_test_alt_vs_ref_full_table)), "adj_combined_pvalue_HSO"] < 0.05), -->
<!--                                     sum(fisher_test_alt_vs_ref_full_table[grep("Middle", rownames(fisher_test_alt_vs_ref_full_table)), "adj_combined_pvalue_HSO"] < 0.05), -->
<!--                                     sum(fisher_test_alt_vs_ref_full_table[grep("Right", rownames(fisher_test_alt_vs_ref_full_table)), "adj_combined_pvalue_HSO"] < 0.05)), -->
<!--                          non_signif = c(sum(fisher_test_alt_vs_ref_full_table[grep("Left", rownames(fisher_test_alt_vs_ref_full_table)), "adj_combined_pvalue_HSO"] > 0.05), -->
<!--                                         sum(fisher_test_alt_vs_ref_full_table[grep("Middle", rownames(fisher_test_alt_vs_ref_full_table)), "adj_combined_pvalue_HSO"] > 0.05), -->
<!--                                         sum(fisher_test_alt_vs_ref_full_table[grep("Right", rownames(fisher_test_alt_vs_ref_full_table)), "adj_combined_pvalue_HSO"] > 0.05))) -->
<!-- rownames(chi_two_df) <- c("Left", "Middle", "Right") -->
<!-- chi_two_df$location = rownames(chi_two_df) -->
<!-- chi_two_df_melt <- melt(chi_two_df) -->
<!-- ggplot(chi_two_df_melt, aes(x = location, y = value, fill = location)) + geom_bar(stat = "identity") +  -->
<!--   facet_wrap(~ variable) + -->
<!--   ggtitle("Barplot of the number of significant combined adjusted p-value\nfor sample HSO on REF sequences") -->
<!-- ``` -->

<!-- #### Analysis on ALT sequences for HSO sample -->

<!-- Here is the boxplot of adjusted combined p-values for sample HSO on ALT sequences. -->

<!-- ```{r} -->
<!-- chi_two_df <- data.frame(signif = c(sum(fisher_test_alt_vs_ref_full_table[grep("Left", rownames(fisher_test_alt_vs_ref_full_table)), "adj_combined_pvalue_HSO"] < 0.05), -->
<!--                                     sum(fisher_test_alt_vs_ref_full_table[grep("Middle", rownames(fisher_test_alt_vs_ref_full_table)), "adj_combined_pvalue_HSO"] < 0.05), -->
<!--                                     sum(fisher_test_alt_vs_ref_full_table[grep("Right", rownames(fisher_test_alt_vs_ref_full_table)), "adj_combined_pvalue_HSO"] < 0.05)), -->
<!--                          non_signif = c(sum(fisher_test_alt_vs_ref_full_table[grep("Left", rownames(fisher_test_alt_vs_ref_full_table)), "adj_combined_pvalue_HSO"] > 0.05), -->
<!--                                         sum(fisher_test_alt_vs_ref_full_table[grep("Middle", rownames(fisher_test_alt_vs_ref_full_table)), "adj_combined_pvalue_HSO"] > 0.05), -->
<!--                                         sum(fisher_test_alt_vs_ref_full_table[grep("Right", rownames(fisher_test_alt_vs_ref_full_table)), "adj_combined_pvalue_HSO"] > 0.05))) -->
<!-- rownames(chi_two_df) <- c("Left", "Middle", "Right") -->
<!-- chi_two_df$location = rownames(chi_two_df) -->
<!-- chi_two_df_melt <- melt(chi_two_df) -->
<!-- ggplot(chi_two_df_melt, aes(x = location, y = value, fill = location)) + geom_bar(stat = "identity") +  -->
<!--   facet_wrap(~ variable) + -->
<!--   ggtitle("Barplot of the number of significant combined adjusted p-value\nfor sample HSO on ALT sequences") -->
<!-- ``` -->

<!-- #### Cochran-Mantel-Haenszel chi-squared test for sample 101 -->

<!-- We can have the feeling that Left shifting has more significant p-values. We can validate this by running a Cochran-Mantel-Haenszel chi-squared test. I run it on the following table. -->

<!-- ```{r} -->
<!-- ftable(chi_two_array) -->

<!-- chi_two_array <- array(c(3256, 691, 3360, 691, -->
<!--                          2984, 963, 3038, 1013, -->
<!--                          2850, 1097, 2872, 1179), -->
<!--                        dim = c(2,2,3), -->
<!--                        dimnames = list(Significant = c("Yes", "No"), -->
<!--                                        Allele = c("REF", "ALT"), -->
<!--                                        Location = c("Left", "Middle", "Right"))) -->
<!-- # mantelhaen.test(chi_two_array) -->
<!-- ``` -->

<!-- We can look at the odds ratio to have a have an idea of the future result -->

<!-- ```{r} -->
<!-- apply(chi_two_array, 3, odds.ratio) -->
<!-- ``` -->

<!-- The odds ratio are very close. We can think that the significance of the Fisher test on REF or ALT sequence is not dependent to the sequence shifting. -->

<!-- And here is the chi-squared results -->

<!-- ```{r} -->
<!-- mantelhaen_results <- mantelhaen.test(chi_two_array) -->
<!-- mantelhaen_results_df <- data.frame(p.value = mantelhaen_results$p.value, odd_ratio = mantelhaen_results$estimate) -->
<!-- rownames(mantelhaen_results_df) <- "" -->
<!-- kable(mantelhaen_results_df) -->
<!-- ``` -->

<!-- The p-value is not significant, as expected above. -->