---
output: html_document
editor_options: 
  chunk_output_type: console
---

## Fisher test to mesure the influence of a sequence versus all other in cDNA and gDNA expression level, splitted by location (Left, Middle and Right)

On the following analysis, we focused on measuring the effect of the sequence versus the other sequence on transcription.

This was perform by creating a contingency matrix of the count of the sequence and the sum of count of all other sequence in the cDNA and the gDNA sample. Contingency matrices look like the following:

|     | cDNA | gDNA |
|-----|------|------|
| rs9999446-ALT-Left-1 | 138   | 74 |
| sum(all_other_sequence) | 8792321   | 5751872 |

This example is for rs9999446-REF-Middle-1 101_replicat_1.

```{r fisher_test_rs_vs_notrs_function}
# Create a function to compute fisher test of sequence effect
# Function to create contingency matrix to perform fisher test
create_cont_matrix_and_run_fisher_test_sequence_vs_not_sequence <- function(sequence_name, samplename, count_mat){
  cont_matrix <- matrix(data = NA, nrow = 2, ncol = 2)
  rownames(cont_matrix) <- c("rs", "!rs")
  colnames(cont_matrix) <- c("cDNA", "gDNA")
  cont_matrix[1,1] <- count_mat[grep(sequence_name, rownames(count_mat)), samplename]
  cont_matrix[2,1] <- sum(count_mat[-grep(sequence_name, rownames(count_mat)), samplename])
  cont_matrix[1,2] <- count_mat[grep(sequence_name, rownames(count_mat)), 
                                gsub(pattern = "cDNA", replacement = "ADNg", x = samplename)]
  cont_matrix[2,2] <- sum(count_mat[-grep(sequence_name, rownames(count_mat)), 
                                    gsub(pattern = "cDNA", replacement = "ADNg", x = samplename)])
  # Make on fisher test and extract value
  fisher_test <- fisher.test(x = cont_matrix, alternative = "two.sided")
  ficher_test_pval <- fisher_test$p.value
  ficher_test_oddratio <- fisher_test$estimate
  return(list(pval = ficher_test_pval, oddratio = ficher_test_oddratio))
}
```

```{r fisher_test_rs_vs_notrs}
# Loop to create contingency table and run the fischer test
fisher_test_rs_vs_notrs <- matrix(data = NA, nrow = length(rownames(count_matrix)), ncol = length(colnames_order_cDNA))
rownames(fisher_test_rs_vs_notrs) <- rownames(count_matrix)
colnames(fisher_test_rs_vs_notrs) <- gsub(pattern = "cDNA_", replacement = "", x = colnames_order_cDNA)
fisher_test_rs_vs_notrs_odd_ratio <- fisher_test_rs_vs_notrs

for(sample in colnames_order_cDNA){
  for(i in 1:nrow(count_matrix)){
    rs_name <- rownames(count_matrix)[i]
    fisher_test_loop <- create_cont_matrix_and_run_fisher_test_sequence_vs_not_sequence(sequence_name = rs_name, 
                                                                                        samplename = sample, 
                                                                                        count_mat = count_matrix)
    fisher_test_rs_vs_notrs[rs_name, gsub(pattern = "_cDNA", replacement = "", 
                                          x = sample)] <- fisher_test_loop$pval
    fisher_test_rs_vs_notrs_odd_ratio[rs_name, gsub(pattern = "_cDNA", replacement = "", 
                                                    x = sample)] <- fisher_test_loop$oddratio    
  }
}
```

After performing the all the Fisher tests, we can draw the following plot colored by "location" (i.e Left/Middle/Right).

On this plot, the lower triangle contains scatter plot of -log10(p-value) paired by samples. The diagonal represent the distribution of the -log10(p-value) and the upper triangle shows the pearson correlation between -log10(p-value) of each sample (gray) and for each sample splitted by "location" (red/green/blue).

### P-value scatter plots between all sample

```{r fisher_test_rs_vs_notrs_plots}
colnames(fisher_test_rs_vs_notrs_odd_ratio) <- paste0("odd_ratio_", colnames(fisher_test_rs_vs_notrs_odd_ratio))

# Replace 0 with a very low value to avoir infinite value when transforming into log10.
fisher_test_rs_vs_notrs[fisher_test_rs_vs_notrs==0] <- 2.737124e-321

# fisher_test_rs_vs_notrs_df <- as.data.frame(fisher_test_rs_vs_notrs)
# fisher_test_rs_vs_notrs_df$location <- gsub(pattern = "rs[0-9]*-|-[123]", replacement = "", x = rownames(fisher_test_rs_vs_notrs_df))
# fisher_test_rs_vs_notrs_df$allele <- gsub(pattern = "rs[0-9]*-|-[123]", replacement = "", x = rownames(fisher_test_rs_vs_notrs_df))
# ggpairs(fisher_test_rs_vs_notrs_df, aes(color = location)) + ggtitle("Plot matrix for raw fisher test p-value")

fisher_test_rs_vs_notrs_df_log <- as.data.frame(-log10(fisher_test_rs_vs_notrs))
# Assign minimal possible value to 0 in the dataframe
# fisher_test_rs_vs_notrs_df_log[is.infinite(fisher_test_rs_vs_notrs_df_log)] <- max(fisher_test_rs_vs_notrs_df_log)
fisher_test_rs_vs_notrs_df_log$location <- gsub(pattern = "rs[0-9]*-|-[123]", replacement = "", x = rownames(fisher_test_rs_vs_notrs_df_log))
# fisher_test_rs_vs_notrs_df_log[fisher_test_rs_vs_notrs_df_log]
# fisher_test_rs_vs_notrs_df_log_only_finite <- fisher_test_rs_vs_notrs_df_log[is.finite(fisher_test_rs_vs_notrs_df_log$`101_replicat_1`)&is.finite(fisher_test_rs_vs_notrs_df_log$`101_replicat_2`)&is.finite(fisher_test_rs_vs_notrs_df_log$HSO_replicat_1)&is.finite(fisher_test_rs_vs_notrs_df_log$HSO_replicat_2),]
# ggplot(fisher_results_df_log, aes (x = HSO_replicat_1, y = HSO_replicat_2, color = location)) + geom_point() + geom_vline(xintercept = 20) + geom_hline(yintercept = 20)

# fisher_results_df
# pairs(fisher_results_df_log[,-grep("location", colnames(fisher_results_df_log))], lower.panel = dens_plotFun, upper.panel = NULL, main = "Fisher test p-value for each condition")
ggpairs(fisher_test_rs_vs_notrs_df_log[,1:4], aes(color = fisher_test_rs_vs_notrs_df_log$location)) + ggtitle("Plot matrix for -log10 fisher test p-value")
```

We observe a better correlation between HTO replicates than for 101 replicates.

### Table of significant p-value

The following table shows the number of p-value lower than 0.01 (i.e log2(p-value) greater than 2). 

```{r}
rep1_101_signif_2 <- sum(fisher_test_rs_vs_notrs_df_log$`101_replicat_1` > 2)
rep2_101_signif_2 <- sum(fisher_test_rs_vs_notrs_df_log$`101_replicat_2` > 2)
rep1_HSO_signif_2 <- sum(fisher_test_rs_vs_notrs_df_log$HSO_replicat_1 > 2)
rep2_HSO_signif_2 <- sum(fisher_test_rs_vs_notrs_df_log$HSO_replicat_2 > 2)
signif_df <- as.data.frame(t(data.frame(replicate1_101 = rep1_101_signif_2,
                                        replicate2_101 = rep2_101_signif_2,
                                        replicate1_HSO = rep1_HSO_signif_2,
                                        replicate2_HSO = rep2_HSO_signif_2)))
control_sequence_nb_rep1_101 <- sum(control_table$rsID. %in% gsub(pattern = "-.*", replacement = "", 
                                                                  x = rownames(fisher_test_rs_vs_notrs_df_log[fisher_test_rs_vs_notrs_df_log$`101_replicat_1` > 2,])))
control_sequence_nb_rep2_101 <- sum(control_table$rsID. %in% gsub(pattern = "-.*", replacement = "", 
                                                                  x = rownames(fisher_test_rs_vs_notrs_df_log[fisher_test_rs_vs_notrs_df_log$`101_replicat_2` > 2,])))
control_sequence_nb_rep1_HSO <- sum(control_table$rsID. %in% gsub(pattern = "-.*", replacement = "", 
                                                                  x = rownames(fisher_test_rs_vs_notrs_df_log[fisher_test_rs_vs_notrs_df_log$HSO_replicat_1 > 2,])))
control_sequence_nb_rep2_HSO <- sum(control_table$rsID. %in% gsub(pattern = "-.*", replacement = "", 
                                                                  x = rownames(fisher_test_rs_vs_notrs_df_log[fisher_test_rs_vs_notrs_df_log$HSO_replicat_2 > 2,])))

colnames(signif_df) <- c("Number_of_p-value_lower_than_0.01")
signif_df$Number_of_significant_control <- c(control_sequence_nb_rep1_101, control_sequence_nb_rep2_101, control_sequence_nb_rep1_HSO, control_sequence_nb_rep2_HSO)
# DT::datatable(signif_df)
kable(signif_df)
# DT::datatable(t(data.frame(replicate1_101 = rep1_101_signif_2,
#                            replicate2_101 = rep2_101_signif_2,
#                            replicate1_HSO = rep1_HSO_signif_2,
#                            replicate2_HSO = rep2_HSO_signif_2)), colnames = "Number of p-value lower than 0.01")
```

### Fisher test p-value for highly significant sequences

After that, we extracted the most significant p-value with a threshold at 1e-200. On the following plot, only sequence with a Fisher p-value lower than 1e-6 are shown. 

```{r getting_most_significant_p-value_seq_vs_notseq}
fisher_results_df_log_highly_significant_all_sample <- fisher_test_rs_vs_notrs_df_log[fisher_test_rs_vs_notrs_df_log$`101_replicat_1` > 200 & fisher_test_rs_vs_notrs_df_log$`101_replicat_2` > 200 & fisher_test_rs_vs_notrs_df_log$HSO_replicat_1 > 200 & fisher_test_rs_vs_notrs_df_log$HSO_replicat_2 > 200,]
# ggpairs(fisher_results_df_log_highly_significant_all_sample[,1:4], aes(color = fisher_results_df_log_highly_significant_all_sample$location)) + ggtitle("Plot matrix for -log10 fisher test p-value")

fisher_results_df_log_highly_significant_all_sample$is_control <- "No"
rs_in_control <- rownames(fisher_results_df_log_highly_significant_all_sample)[gsub(pattern = "-.*", replacement = "", x = rownames(fisher_results_df_log_highly_significant_all_sample)) %in% control_table$rsID.]
fisher_results_df_log_highly_significant_all_sample[rs_in_control, "is_control"] <- "Yes"
fisher_results_df_log_highly_significant_all_sample$name <- rownames(fisher_results_df_log_highly_significant_all_sample)

# ggpairs(fisher_results_df_log_highly_significant_all_sample[, 1:4], aes(color = as.factor(fisher_results_df_log_highly_significant_all_sample$is_control)), lower = list(continuous = function(data, mapping, ...) ggally_points(data, mapping, ...) + ggrepel::geom_label_repel(aes(label = fisher_results_df_log_highly_significant_all_sample$name), max.overlaps = 50)), upper = NULL) + ggtitle("Plot matrix for -log10 p-value significant in all samples") +   
#   theme(legend.position = "right") + 
#   labs(fill = "is_control")

fisher_results_df_log_highly_significant_all_sample %>% ggpairs(., legend = 1, columns = 1:4, 
                                                                aes(color = fisher_results_df_log_highly_significant_all_sample$is_control), lower = list(continuous = function(data, mapping, ...) ggally_points(data, mapping, ...) + ggrepel::geom_label_repel(aes(label = fisher_results_df_log_highly_significant_all_sample$name), max.overlaps = 50)), upper = NULL) + 
  theme(legend.position = "bottom") + 
  labs(fill = "is_control")
```

```{r}
fisher_test_rs_vs_notrs_samples <- as.data.frame(fisher_test_rs_vs_notrs[,grep("rep", colnames(fisher_test_rs_vs_notrs))])

colnames(fisher_test_rs_vs_notrs_samples) <- paste0("pvalue_", colnames(fisher_test_rs_vs_notrs_samples))
colnames(fisher_test_rs_vs_notrs_odd_ratio) <- paste0("odd_ratio_", colnames(fisher_test_rs_vs_notrs))

# Correct p-value using the BH method
# Use the index contained in the results of mt.rawp2adjp to get the location of the sequence in the matrix.
adjusted_pval <- apply(fisher_test_rs_vs_notrs_samples, 2, mt.rawp2adjp, proc="BH")
fisher_test_rs_vs_notrs_samples$adj_pvalue_101_replicat_1 <- adjusted_pval$pvalue_101_replicat_1$adjp[order(adjusted_pval$pvalue_101_replicat_1$index),][,2]
fisher_test_rs_vs_notrs_samples$adj_pvalue_101_replicat_2 <- adjusted_pval$pvalue_101_replicat_2$adjp[order(adjusted_pval$pvalue_101_replicat_2$index),][,2]
fisher_test_rs_vs_notrs_samples$adj_pvalue_HSO_replicat_1 <- adjusted_pval$pvalue_HSO_replicat_1$adjp[order(adjusted_pval$pvalue_HSO_replicat_1$index),][,2]
fisher_test_rs_vs_notrs_samples$adj_pvalue_HSO_replicat_2 <- adjusted_pval$pvalue_HSO_replicat_2$adjp[order(adjusted_pval$pvalue_HSO_replicat_2$index),][,2]

# Combine fisher test p-value 
# rv_vs_notrs_101_combined_pval <- apply(fisher_test_rs_vs_notrs_samples[,grep("adj_pvalue_101", colnames(fisher_test_rs_vs_notrs_samples))], 1, sumlog)
# fisher_test_rs_vs_notrs_samples <- cbind(fisher_test_rs_vs_notrs_samples, combined_pval_101 = do.call(rbind,lapply(rv_vs_notrs_101_combined_pval,function(v){v$p})))
# rv_vs_notrs_HSO_combined_pval <- apply(fisher_test_rs_vs_notrs_samples[,grep("adj_pvalue_HSO", colnames(fisher_test_rs_vs_notrs_samples))], 1, sumlog)
# fisher_test_rs_vs_notrs_samples <- cbind(fisher_test_rs_vs_notrs_samples, combined_pval_HSO = do.call(rbind,lapply(rv_vs_notrs_HSO_combined_pval,function(v){v$p})))
rv_vs_notrs_101_combined_pval <- apply(fisher_test_rs_vs_notrs_samples[,grep("pvalue_101_replicat", colnames(fisher_test_rs_vs_notrs_samples))], 1, sumlog)
fisher_test_rs_vs_notrs_samples <- cbind(fisher_test_rs_vs_notrs_samples, combined_pval_101 = do.call(rbind,lapply(rv_vs_notrs_101_combined_pval,function(v){v$p})))
rv_vs_notrs_HSO_combined_pval <- apply(fisher_test_rs_vs_notrs_samples[,grep("pvalue_HSO_replicat", colnames(fisher_test_rs_vs_notrs_samples))], 1, sumlog)
fisher_test_rs_vs_notrs_samples <- cbind(fisher_test_rs_vs_notrs_samples, combined_pval_HSO = do.call(rbind,lapply(rv_vs_notrs_HSO_combined_pval,function(v){v$p})))

adjusted_pval <- apply(fisher_test_rs_vs_notrs_samples[, grep("combined_pval", colnames(fisher_test_rs_vs_notrs_samples))], 2, mt.rawp2adjp, proc="BH")
fisher_test_rs_vs_notrs_samples$adj_combined_pvalue_101 <- adjusted_pval$combined_pval_101$adjp[order(adjusted_pval$combined_pval_101$index),][,2]
fisher_test_rs_vs_notrs_samples$adj_combined_pvalue_HSO <- adjusted_pval$combined_pval_HSO$adjp[order(adjusted_pval$combined_pval_HSO$index),][,2]

fisher_test_rs_vs_notrs_full_table <- cbind(count_matrix, fisher_test_rs_vs_notrs_samples, fisher_test_rs_vs_notrs_odd_ratio)
# Check if the odd ratio are concordant between replicates
fisher_test_rs_vs_notrs_full_table$concordance_odd_ratio_101 <- FALSE
fisher_test_rs_vs_notrs_full_table[(fisher_test_rs_vs_notrs_full_table$odd_ratio_101_replicat_1 < 1 &
                                      fisher_test_rs_vs_notrs_full_table$odd_ratio_101_replicat_2 < 1) | 
                                     (fisher_test_rs_vs_notrs_full_table$odd_ratio_101_replicat_1 > 1 & 
                                        fisher_test_rs_vs_notrs_full_table$odd_ratio_101_replicat_2 > 1) | 
                                     (fisher_test_rs_vs_notrs_full_table$odd_ratio_101_replicat_1 == 1 &
                                        fisher_test_rs_vs_notrs_full_table$odd_ratio_101_replicat_2 == 1), "concordance_odd_ratio_101"] <- TRUE
fisher_test_rs_vs_notrs_full_table$concordance_odd_ratio_HSO <- FALSE
fisher_test_rs_vs_notrs_full_table[(fisher_test_rs_vs_notrs_full_table$odd_ratio_HSO_replicat_1 < 1 &
                                      fisher_test_rs_vs_notrs_full_table$odd_ratio_HSO_replicat_2 < 1) | 
                                     (fisher_test_rs_vs_notrs_full_table$odd_ratio_HSO_replicat_1 > 1 & 
                                        fisher_test_rs_vs_notrs_full_table$odd_ratio_HSO_replicat_2 > 1) | 
                                     (fisher_test_rs_vs_notrs_full_table$odd_ratio_HSO_replicat_1 == 1 &
                                        fisher_test_rs_vs_notrs_full_table$odd_ratio_HSO_replicat_2 == 1), "concordance_odd_ratio_HSO"] <- TRUE

fisher_test_rs_vs_notrs_full_table$mean_odd_ratio_101 <- log2(rowMeans(fisher_test_rs_vs_notrs_full_table[, grep("odd_ratio_101", colnames(fisher_test_rs_vs_notrs_full_table))]))
fisher_test_rs_vs_notrs_full_table$mean_odd_ratio_HSO <- log2(rowMeans(fisher_test_rs_vs_notrs_full_table[, grep("odd_ratio_HSO", colnames(fisher_test_rs_vs_notrs_full_table))]))
fisher_test_rs_vs_notrs_full_table$concordance_odd_ratio_101 <- as.factor(fisher_test_rs_vs_notrs_full_table$concordance_odd_ratio_101)
fisher_test_rs_vs_notrs_full_table$concordance_odd_ratio_HSO <- as.factor(fisher_test_rs_vs_notrs_full_table$concordance_odd_ratio_HSO)

fisher_test_rs_vs_notrs_full_table$qval_adj_combined_pvalue_101 <- qvalue(fisher_test_rs_vs_notrs_full_table$adj_combined_pvalue_101)$qvalues
fisher_test_rs_vs_notrs_full_table$qval_adj_combined_pvalue_101_FDR_5 <- as.factor(qvalue(fisher_test_rs_vs_notrs_full_table$adj_combined_pvalue_101, fdr.level = 0.05)$significant)
fisher_test_rs_vs_notrs_full_table$qval_adj_combined_pvalue_101_FDR_10 <- as.factor(qvalue(fisher_test_rs_vs_notrs_full_table$adj_combined_pvalue_101, fdr.level = 0.10)$significant)
fisher_test_rs_vs_notrs_full_table$qval_adj_combined_pvalue_HSO <- qvalue(fisher_test_rs_vs_notrs_full_table$adj_combined_pvalue_HSO)$qvalues
fisher_test_rs_vs_notrs_full_table$qval_adj_combined_pvalue_HSO_FDR_5 <- as.factor(qvalue(fisher_test_rs_vs_notrs_full_table$adj_combined_pvalue_HSO, fdr.level = 0.05)$significant)
fisher_test_rs_vs_notrs_full_table$qval_adj_combined_pvalue_HSO_FDR_10 <- as.factor(qvalue(fisher_test_rs_vs_notrs_full_table$adj_combined_pvalue_HSO, fdr.level = 0.10)$significant)
```

### Volcano plot

The following volcano plots show the relation between the p-value and the odd ratio. The plots are splitted depending on the concordance of odd ratio. TRUE indicates concordants odd ratio, whereas FALSE indicante non concordant ones.

```{r volcano_qval_oddratio_rs_vs_notrs}
ggplot(fisher_test_rs_vs_notrs_full_table, aes(x = mean_odd_ratio_101, y = -log10(adj_combined_pvalue_101), color = concordance_odd_ratio_101)) + 
  geom_point() + 
  facet_wrap(~ concordance_odd_ratio_101) +
  ggtitle("Volcano plot of the -log10 of adjusted combined p-value for samples 101") + 
  ylab("-log10(adjusted combined p-value of Fisher test)") +
  xlab("Mean odd ratio of Fisher test") + 
  labs(color = "Are odd ratio concordant") +
  scale_colour_discrete(labels = paste(levels(fisher_test_rs_vs_notrs_full_table$concordance_odd_ratio_101), table(fisher_test_rs_vs_notrs_full_table$concordance_odd_ratio_101)))

ggplot(fisher_test_rs_vs_notrs_full_table, aes(x = mean_odd_ratio_HSO, y = -log10(adj_combined_pvalue_HSO), color = concordance_odd_ratio_HSO)) + 
  geom_point() + 
  facet_wrap(~ concordance_odd_ratio_HSO) +
  ggtitle("Volcano plot f the -log10 of adjusted combined p-value for samples HSO") + 
  ylab("-log10(adjusted combined p-value of Fisher test)") +
  xlab("Mean odd ratio of Fisher test") + 
  labs(color = "Are odd ratio concordant") +
  scale_colour_discrete(labels = paste(levels(fisher_test_rs_vs_notrs_full_table$concordance_odd_ratio_HSO), table(fisher_test_rs_vs_notrs_full_table$concordance_odd_ratio_HSO)))
```

### Export table

I created and exported a table containing all results for this pool of Fisher tests. The table contains:

* Raw counts of each sequence for each samples
* p-value of Fisher test of each sequence for each sample
* Combined p-value using the Fisher method
* Adjusted p-value of Fisher test using the Benjamini-Hochberg correction method for multiple testing
* Combined adjusted p-value using the Fisher method
* The odd ratio of Fisher test for each sequence and each sample
* A boolean column (TRUE of FALSE) for each sample indicating if odd ratio for both replicate are concordant (both < 1 or both = 1 or both > 1)
* The mean odd ratio of each sample
* Qvalue of combined adjusted p-value
* A boolean column identifying significant sequences with a FDR of 5% or 10% (TRUE or FALSE).

```{r}
write.csv(x = fisher_test_rs_vs_notrs_full_table,
          file = "/data/nin/Workspace/Runs/Run_384/Tables/Table_Fisher_tests_chapter_2.2.csv", 
          row.names = TRUE)
```

### Table with some sequences having a significant regulatory effect

The following table shows sequences having a significant regulatory effect. The shown sequences are selected by having an adjusted combined p-value (using the Fisher combination method for p-value) lower than 0.05

#### Sample 101

```{r}
sample_101_significant_seq_combined_pval <- as.data.frame( rownames(fisher_test_rs_vs_notrs_full_table[fisher_test_rs_vs_notrs_full_table$adj_combined_pvalue_101 < 0.05,]))
colnames(sample_101_significant_seq_combined_pval) <- "significant_sequences"
datatable(sample_101_significant_seq_combined_pval)
```

#### Sample HSO

```{r}
sample_HSO_significant_seq_combined_pval <- as.data.frame( rownames(fisher_test_rs_vs_notrs_full_table[fisher_test_rs_vs_notrs_full_table$adj_combined_pvalue_HSO < 0.05,]))
colnames(sample_HSO_significant_seq_combined_pval) <- "significant_sequences"
datatable(sample_HSO_significant_seq_combined_pval)
```

#### Common significant sequences

The following venn diagram shows the number of significant sequences detected in both samples.

```{r}
# Fonction d'aide pour afficher le diagramme de Venn
display_venn <- function(x, ...){
  library(VennDiagram)
  grid.newpage()
  venn_object <- venn.diagram(x, filename = NULL, ...)
  grid.draw(venn_object)
}
display_venn(
  x = c(sample_HSO_significant_seq_combined_pval, sample_101_significant_seq_combined_pval),
  category.names = c(paste0("HSO\nn = ", nrow(sample_HSO_significant_seq_combined_pval)), 
                     paste0("101\nn = ", nrow(sample_101_significant_seq_combined_pval))),
  # Cercles
  lwd = 2,
  lty = 'blank',
  fill = c("#999999", "#E69F00"),
  # Nombres
  cex = .9,
  fontface = "italic",
  # Noms des groupes
  cat.cex = 1,
  cat.fontface = "bold",
  cat.default.pos = "outer",
  cat.dist = c(0.055, 0.055)
)
```

Here is the list of common detected sequences

```{r}
common_significant_seq_combined_pval <- as.data.frame(intersect(sample_101_significant_seq_combined_pval$significant_sequences, sample_HSO_significant_seq_combined_pval$significant_sequences))
colnames(common_significant_seq_combined_pval) <- "Common_significant_sequences"
datatable(common_significant_seq_combined_pval)
```

### Qvalues

I computed the qvalues of adjusted combined p-values for each samples. I draw the following histogram for each sample. I shows the distribution of adjusted combined p-value and draw a line for the qvalues and the local FDR. I also draw 2 volcano plots showing the relation between the p-value and the odd ratio. The plots are splitted depending on the significance of the qvalues. TRUE indicates a significant qvalue, whereas FALSE indicates non significant ones. I draw the plots using a FDR of 5% and 10%.

#### Sample 101

```{r}
# hist(qvalue(fisher_test_rs_vs_notrs_full_table$adj_combined_pvalue_101))

ggplot(fisher_test_rs_vs_notrs_full_table, aes(x = mean_odd_ratio_101, y = -log10(adj_combined_pvalue_101), color = qval_adj_combined_pvalue_101_FDR_5)) + 
  geom_point() + 
  facet_wrap(~ qval_adj_combined_pvalue_101_FDR_5) +
  ggtitle("Volcano plot of the -log10 of adjusted combined p-value for samples 101") + 
  ylab("-log10(adjusted combined p-value of Fisher test)") +
  xlab("Mean odd ratio of Fisher test") + 
  labs(color = "Is qvalue significant at FDR 5%?") +
  scale_colour_discrete(labels = paste(levels(fisher_test_rs_vs_notrs_full_table$qval_adj_combined_pvalue_101_FDR_5), table(fisher_test_rs_vs_notrs_full_table$qval_adj_combined_pvalue_101_FDR_5)))

ggplot(fisher_test_rs_vs_notrs_full_table, aes(x = mean_odd_ratio_101, y = -log10(adj_combined_pvalue_101), color = qval_adj_combined_pvalue_101_FDR_10)) + 
  geom_point() + 
  facet_wrap(~ qval_adj_combined_pvalue_101_FDR_10) +
  ggtitle("Volcano plot of the -log10 of adjusted combined p-value for samples 101") + 
  ylab("-log10(adjusted combined p-value of Fisher test)") +
  xlab("Mean odd ratio of Fisher test") + 
  labs(color = "Is qvalue significant at FDR 10%?") +
  scale_colour_discrete(labels = paste(levels(fisher_test_rs_vs_notrs_full_table$qval_adj_combined_pvalue_101_FDR_10), table(fisher_test_rs_vs_notrs_full_table$qval_adj_combined_pvalue_101_FDR_10)))
```

##### List of significant sequence with a 5% FDR

```{r}
datatable(fisher_test_rs_vs_notrs_full_table[fisher_test_rs_vs_notrs_full_table$qval_adj_combined_pvalue_101_FDR_5 == TRUE, 
                                             c("adj_combined_pvalue_101", "qval_adj_combined_pvalue_101")])
```

##### List of significant sequence with a 10% FDR

```{r}
datatable(fisher_test_rs_vs_notrs_full_table[fisher_test_rs_vs_notrs_full_table$qval_adj_combined_pvalue_101_FDR_10 == TRUE, 
                                             c("adj_combined_pvalue_101", "qval_adj_combined_pvalue_101")])
```

#### Sample HSO

```{r}
# hist(qvalue(fisher_test_rs_vs_notrs_full_table$adj_combined_pvalue_HSO))

ggplot(fisher_test_rs_vs_notrs_full_table, aes(x = mean_odd_ratio_HSO, y = -log10(adj_combined_pvalue_HSO), color = qval_adj_combined_pvalue_HSO_FDR_5)) + 
  geom_point() + 
  facet_wrap(~ qval_adj_combined_pvalue_HSO_FDR_5) +
  ggtitle("Volcano plot of the -log10 of adjusted combined p-value for samples HSO") + 
  ylab("-log10(adjusted combined p-value of Fisher test)") +
  xlab("Mean odd ratio of Fisher test") + 
  labs(color = "Is qvalue significant at FDR 5%?") +
  scale_colour_discrete(labels = paste(levels(fisher_test_rs_vs_notrs_full_table$qval_adj_combined_pvalue_HSO_FDR_5), table(fisher_test_rs_vs_notrs_full_table$qval_adj_combined_pvalue_HSO_FDR_5)))

ggplot(fisher_test_rs_vs_notrs_full_table, aes(x = mean_odd_ratio_HSO, y = -log10(adj_combined_pvalue_HSO), color = qval_adj_combined_pvalue_HSO_FDR_10)) + 
  geom_point() + 
  facet_wrap(~ qval_adj_combined_pvalue_HSO_FDR_10) +
  ggtitle("Volcano plot of the -log10 of adjusted combined p-value for samples HSO") + 
  ylab("-log10(adjusted combined p-value of Fisher test)") +
  xlab("Mean odd ratio of Fisher test") + 
  labs(color = "Is qvalue significant at FDR 10%?") +
  scale_colour_discrete(labels = paste(levels(fisher_test_rs_vs_notrs_full_table$qval_adj_combined_pvalue_HSO_FDR_10), table(fisher_test_rs_vs_notrs_full_table$qval_adj_combined_pvalue_HSO_FDR_10)))
```

##### List of significant sequence with a 5% FDR

```{r}
datatable(fisher_test_rs_vs_notrs_full_table[fisher_test_rs_vs_notrs_full_table$qval_adj_combined_pvalue_HSO_FDR_5 == TRUE, 
                                             c("adj_combined_pvalue_HSO", "qval_adj_combined_pvalue_HSO")])
```

##### List of significant sequence with a 10% FDR

```{r}
datatable(fisher_test_rs_vs_notrs_full_table[fisher_test_rs_vs_notrs_full_table$qval_adj_combined_pvalue_HSO_FDR_10 == TRUE, 
                                             c("adj_combined_pvalue_HSO", "qval_adj_combined_pvalue_HSO")])
```

#### Venn diagrams with qvalues

##### FDR 5%

```{r}
sample_101_significant_seq_qval <- as.data.frame( rownames(fisher_test_rs_vs_notrs_full_table[fisher_test_rs_vs_notrs_full_table$qval_adj_combined_pvalue_101_FDR_5 == TRUE, 
                                                                                              c("adj_combined_pvalue_101", "qval_adj_combined_pvalue_101")]))
colnames(sample_101_significant_seq_qval) <- "significant_sequences_with_FDR_5"

sample_HSO_significant_seq_qval <- as.data.frame( rownames(fisher_test_rs_vs_notrs_full_table[fisher_test_rs_vs_notrs_full_table$qval_adj_combined_pvalue_HSO_FDR_5 == TRUE, 
                                                                                              c("adj_combined_pvalue_HSO", "qval_adj_combined_pvalue_HSO")]))
colnames(sample_HSO_significant_seq_qval) <- "significant_sequences_with_FDR_5"

display_venn(
  x = c(sample_HSO_significant_seq_qval, sample_101_significant_seq_qval),
  category.names = c(paste0("HSO\nn = ", nrow(sample_HSO_significant_seq_qval)), 
                     paste0("101\nn = ", nrow(sample_101_significant_seq_qval))),
  # Cercles
  lwd = 2,
  lty = 'blank',
  fill = c("#999999", "#E69F00"),
  # Nombres
  cex = .9,
  fontface = "italic",
  # Noms des groupes
  cat.cex = 1,
  cat.fontface = "bold",
  cat.default.pos = "outer",
  cat.dist = c(0.055, 0.055)
)
```

Here is the list of common detected sequences

```{r}
common_significant_seq_combined_pval <- as.data.frame(intersect(sample_HSO_significant_seq_qval$significant_sequences_with_FDR_5, sample_101_significant_seq_qval$significant_sequences_with_FDR_5))
colnames(common_significant_seq_combined_pval) <- "Common_significant_sequences_with_FDR_5%"
datatable(common_significant_seq_combined_pval)
```

##### FDR 10%

```{r}
sample_101_significant_seq_qval <- as.data.frame( rownames(fisher_test_rs_vs_notrs_full_table[fisher_test_rs_vs_notrs_full_table$qval_adj_combined_pvalue_101_FDR_10 == TRUE, 
                                                                                              c("adj_combined_pvalue_101", "qval_adj_combined_pvalue_101")]))
colnames(sample_101_significant_seq_qval) <- "significant_sequences_with_FDR_10"

sample_HSO_significant_seq_qval <- as.data.frame( rownames(fisher_test_rs_vs_notrs_full_table[fisher_test_rs_vs_notrs_full_table$qval_adj_combined_pvalue_HSO_FDR_10 == TRUE, 
                                                                                              c("adj_combined_pvalue_HSO", "qval_adj_combined_pvalue_HSO")]))
colnames(sample_HSO_significant_seq_qval) <- "significant_sequences_with_FDR_10"

display_venn(
  x = c(sample_HSO_significant_seq_qval, sample_101_significant_seq_qval),
  category.names = c(paste0("HSO\nn = ", nrow(sample_HSO_significant_seq_qval)), 
                     paste0("101\nn = ", nrow(sample_101_significant_seq_qval))),
  # Cercles
  lwd = 2,
  lty = 'blank',
  fill = c("#999999", "#E69F00"),
  # Nombres
  cex = .9,
  fontface = "italic",
  # Noms des groupes
  cat.cex = 1,
  cat.fontface = "bold",
  cat.default.pos = "outer",
  cat.dist = c(0.055, 0.055)
)
```

Here is the list of common detected sequences

```{r}
common_significant_seq_combined_pval <- as.data.frame(intersect(sample_HSO_significant_seq_qval$significant_sequences_with_FDR_10, sample_101_significant_seq_qval$significant_sequences_with_FDR_10))
colnames(common_significant_seq_combined_pval) <- "Common_significant_sequences_with_FDR_5%"
datatable(common_significant_seq_combined_pval)
```

### Analysis of the impact of sequence shifting

The following analysis aims to detect an impact of the sequence shifting on the transcription. 

For both samples, (101 and HSO), I looked at the adjusted combined p-values of Fisher test obtained when looking at the impact on transcription for each sequences. Before going into statistical tests, I draw some boxplot of p-value on each samples and separated by significance of p-value (i.e adjusted combined p-values lower than 0.05). 

#### Analysis on REF sequences for 101 sample

Here is the boxplot of adjusted combined p-values for sample 101 on REF sequences.

```{r}
# left_signif <- fisher_test_rs_vs_notrs_full_table[grep("REF-Left", rownames(fisher_test_rs_vs_notrs_full_table[fisher_test_rs_vs_notrs_full_table$qval_adj_combined_pvalue_101 < 0.05,])),]

ref_left_signif <- sum(fisher_test_rs_vs_notrs_full_table[grep("REF-Left", rownames(fisher_test_rs_vs_notrs_full_table)), "adj_combined_pvalue_101"] < 0.05)
ref_left_nosignif <- sum(fisher_test_rs_vs_notrs_full_table[grep("REF-Left", rownames(fisher_test_rs_vs_notrs_full_table)), "adj_combined_pvalue_101"] > 0.05)
alt_left_signif <- sum(fisher_test_rs_vs_notrs_full_table[grep("ALT-Left", rownames(fisher_test_rs_vs_notrs_full_table)), "adj_combined_pvalue_101"] < 0.05)
alt_left_nosignif <- sum(fisher_test_rs_vs_notrs_full_table[grep("ALT-Left", rownames(fisher_test_rs_vs_notrs_full_table)), "adj_combined_pvalue_101"] > 0.05)
ref_middle_signif <- sum(fisher_test_rs_vs_notrs_full_table[grep("REF-Middle", rownames(fisher_test_rs_vs_notrs_full_table)), "adj_combined_pvalue_101"] < 0.05)
ref_middle_nosignif <- sum(fisher_test_rs_vs_notrs_full_table[grep("REF-Middle", rownames(fisher_test_rs_vs_notrs_full_table)), "adj_combined_pvalue_101"] > 0.05)
alt_middle_signif <- sum(fisher_test_rs_vs_notrs_full_table[grep("ALT-Middle", rownames(fisher_test_rs_vs_notrs_full_table)), "adj_combined_pvalue_101"] < 0.05)
alt_middle_nosignif <- sum(fisher_test_rs_vs_notrs_full_table[grep("ALT-Middle", rownames(fisher_test_rs_vs_notrs_full_table)), "adj_combined_pvalue_101"] > 0.05)
ref_right_signif <- sum(fisher_test_rs_vs_notrs_full_table[grep("REF-Right", rownames(fisher_test_rs_vs_notrs_full_table)), "adj_combined_pvalue_101"] < 0.05)
ref_right_nosignif <- sum(fisher_test_rs_vs_notrs_full_table[grep("REF-Right", rownames(fisher_test_rs_vs_notrs_full_table)), "adj_combined_pvalue_101"] > 0.05)
alt_right_signif <- sum(fisher_test_rs_vs_notrs_full_table[grep("ALT-Right", rownames(fisher_test_rs_vs_notrs_full_table)), "adj_combined_pvalue_101"] < 0.05)
alt_right_nosignif <- sum(fisher_test_rs_vs_notrs_full_table[grep("ALT-Right", rownames(fisher_test_rs_vs_notrs_full_table)), "adj_combined_pvalue_101"] > 0.05)

chi_two_array <- array(c(ref_left_signif, ref_left_nosignif, alt_left_signif, alt_left_nosignif,
                         ref_middle_signif, ref_middle_nosignif, alt_middle_signif, alt_middle_nosignif,
                         ref_right_signif, ref_right_nosignif, alt_right_signif, alt_right_nosignif),
                       dim = c(2,2,3),
                       dimnames = list(Significant = c("Yes", "No"),
                                       Allele = c("REF", "ALT"),
                                       Location = c("Left", "Middle", "Right")))
# marginal_table <- margin.table(chi_two_array, c(2,3))
# chi_two_array <- array(c(sum(fisher_test_rs_vs_notrs_full_table[grep("REF-Left", rownames(fisher_test_rs_vs_notrs_full_table)), "adj_combined_pvalue_101"] < 0.05),
#                          sum(fisher_test_rs_vs_notrs_full_table[grep("REF-Left", rownames(fisher_test_rs_vs_notrs_full_table)), "adj_combined_pvalue_101"] > 0.05),
#                          sum(fisher_test_rs_vs_notrs_full_table[grep("REF-Middle", rownames(fisher_test_rs_vs_notrs_full_table)), "adj_combined_pvalue_101"] < 0.05),
#                          sum(fisher_test_rs_vs_notrs_full_table[grep("REF-Middle", rownames(fisher_test_rs_vs_notrs_full_table)), "adj_combined_pvalue_101"] > 0.05),
#                          sum(fisher_test_rs_vs_notrs_full_table[grep("REF-Right", rownames(fisher_test_rs_vs_notrs_full_table)), "adj_combined_pvalue_101"] < 0.05),
#                          sum(fisher_test_rs_vs_notrs_full_table[grep("REF-Right", rownames(fisher_test_rs_vs_notrs_full_table)), "adj_combined_pvalue_101"] > 0.05)),
#                        dim = c(2,3),
#                        dimnames = list(Significant = c("Yes", "No"),
#                                        Location = c("Left", "Middle", "Right")))

# apply(chi_two_array, 3, odds.ratio)
# mantelhaen.test(chi_two_array)
# ftable(chi_two_array)
# 
# chi_two_array <- array(c(2481, 1466, 2483, 1568,
#                          2026, 1921, 2069, 1982,
#                          1761, 2186, 1794, 2257),
#                        dim = c(2,2,3),
#                        dimnames = list(Significant = c("Yes", "No"),
#                                        Allele = c("REF", "ALT"),
#                                        Location = c("Left", "Middle", "Right")))
# mantelhaen.test(chi_two_array)
```

```{r}
# chi_two_array <- array(c(sum(fisher_test_rs_vs_notrs_full_table[grep("REF-Left", rownames(fisher_test_rs_vs_notrs_full_table)), "adj_combined_pvalue_101"] < 0.05),
#                          sum(fisher_test_rs_vs_notrs_full_table[grep("REF-Middle", rownames(fisher_test_rs_vs_notrs_full_table)), "adj_combined_pvalue_101"] < 0.05),
#                          sum(fisher_test_rs_vs_notrs_full_table[grep("REF-Right", rownames(fisher_test_rs_vs_notrs_full_table)), "adj_combined_pvalue_101"] < 0.05),
#                          sum(fisher_test_rs_vs_notrs_full_table[grep("REF-Left", rownames(fisher_test_rs_vs_notrs_full_table)), "adj_combined_pvalue_101"] > 0.05),
#                          sum(fisher_test_rs_vs_notrs_full_table[grep("REF-Middle", rownames(fisher_test_rs_vs_notrs_full_table)), "adj_combined_pvalue_101"] > 0.05),
#                          sum(fisher_test_rs_vs_notrs_full_table[grep("REF-Right", rownames(fisher_test_rs_vs_notrs_full_table)), "adj_combined_pvalue_101"] > 0.05)),
#                        dim = c(2,3),
#                        dimnames = list(Significant = c("Yes", "No"),
#                                        Location = c("Left", "Middle", "Right")))
# 
chi_two_df <- data.frame(signif = c(sum(fisher_test_rs_vs_notrs_full_table[grep("REF-Left", rownames(fisher_test_rs_vs_notrs_full_table)), "adj_combined_pvalue_101"] < 0.05),
                                    sum(fisher_test_rs_vs_notrs_full_table[grep("REF-Middle", rownames(fisher_test_rs_vs_notrs_full_table)), "adj_combined_pvalue_101"] < 0.05),
                                    sum(fisher_test_rs_vs_notrs_full_table[grep("REF-Right", rownames(fisher_test_rs_vs_notrs_full_table)), "adj_combined_pvalue_101"] < 0.05)),
                         non_signif = c(sum(fisher_test_rs_vs_notrs_full_table[grep("REF-Left", rownames(fisher_test_rs_vs_notrs_full_table)), "adj_combined_pvalue_101"] > 0.05),
                                        sum(fisher_test_rs_vs_notrs_full_table[grep("REF-Middle", rownames(fisher_test_rs_vs_notrs_full_table)), "adj_combined_pvalue_101"] > 0.05),
                                        sum(fisher_test_rs_vs_notrs_full_table[grep("REF-Right", rownames(fisher_test_rs_vs_notrs_full_table)), "adj_combined_pvalue_101"] > 0.05)))
rownames(chi_two_df) <- c("Left", "Middle", "Right")
chi_two_df$location = rownames(chi_two_df)
chi_two_df_melt <- melt(chi_two_df)
ggplot(chi_two_df_melt, aes(x = location, y = value, fill = location)) + geom_bar(stat = "identity") + 
  facet_wrap(~ variable) +
  ggtitle("Barplot of the number of significant adjusted combined p-value\nfor sample 101 on REF sequences") 
# mantelhaen.test(chi_two_df)
```

#### Analysis on ALT sequences for 101 sample

Here is the boxplot of adjusted combined p-values for sample 101 on ALT sequences.

```{r}
chi_two_df <- data.frame(signif = c(sum(fisher_test_rs_vs_notrs_full_table[grep("ALT-Left", rownames(fisher_test_rs_vs_notrs_full_table)), "adj_combined_pvalue_101"] < 0.05),
                                    sum(fisher_test_rs_vs_notrs_full_table[grep("ALT-Middle", rownames(fisher_test_rs_vs_notrs_full_table)), "adj_combined_pvalue_101"] < 0.05),
                                    sum(fisher_test_rs_vs_notrs_full_table[grep("ALT-Right", rownames(fisher_test_rs_vs_notrs_full_table)), "adj_combined_pvalue_101"] < 0.05)),
                         non_signif = c(sum(fisher_test_rs_vs_notrs_full_table[grep("ALT-Left", rownames(fisher_test_rs_vs_notrs_full_table)), "adj_combined_pvalue_101"] > 0.05),
                                        sum(fisher_test_rs_vs_notrs_full_table[grep("ALT-Middle", rownames(fisher_test_rs_vs_notrs_full_table)), "adj_combined_pvalue_101"] > 0.05),
                                        sum(fisher_test_rs_vs_notrs_full_table[grep("ALT-Right", rownames(fisher_test_rs_vs_notrs_full_table)), "adj_combined_pvalue_101"] > 0.05)))
rownames(chi_two_df) <- c("Left", "Middle", "Right")
chi_two_df$location = rownames(chi_two_df)
chi_two_df_melt <- melt(chi_two_df)
ggplot(chi_two_df_melt, aes(x = location, y = value, fill = location)) + geom_bar(stat = "identity") + 
  facet_wrap(~ variable) +
  ggtitle("Barplot of the number of significant adjusted combined p-value\nfor sample 101 on ALT sequences")
```

#### Cochran-Mantel-Haenszel chi-squared test for sample 101

We can have the feeling that Left shifting has more significant p-values. We can validate this by running a Cochran-Mantel-Haenszel chi-squared test. I run it on the following table.

```{r}
ftable(chi_two_array)

chi_two_array <- array(c(2481, 1466, 2483, 1568,
                         2026, 1921, 2069, 1982,
                         1761, 2186, 1794, 2257),
                       dim = c(2,2,3),
                       dimnames = list(Significant = c("Yes", "No"),
                                       Allele = c("REF", "ALT"),
                                       Location = c("Left", "Middle", "Right")))
```

We can look at the odds ratio to have a have an idea of the future result

```{r}
apply(chi_two_array, 3, odds.ratio)
```

The odds ratio are very close. We can think that the significance of the Fisher test on REF or ALT sequence is not dependent to the sequence shifting.

And here is the chi-squared results

```{r}
mantelhaen_results <- mantelhaen.test(chi_two_array)
mantelhaen_results_df <- data.frame(p.value = mantelhaen_results$p.value, odd_ratio = mantelhaen_results$estimate)
rownames(mantelhaen_results_df) <- ""
kable(mantelhaen_results_df)
```

The p-value is not significant, as expected above.

#### Analysis on REF sequences for HSO sample

Here is the boxplot of adjusted combined p-values for sample HSO on REF sequences.

```{r}
chi_two_array <- array(c(sum(fisher_test_rs_vs_notrs_full_table[grep("REF-Left", rownames(fisher_test_rs_vs_notrs_full_table)), "adj_combined_pvalue_HSO"] < 0.05),
                         sum(fisher_test_rs_vs_notrs_full_table[grep("REF-Left", rownames(fisher_test_rs_vs_notrs_full_table)), "adj_combined_pvalue_HSO"] > 0.05),
                         sum(fisher_test_rs_vs_notrs_full_table[grep("ALT-Left", rownames(fisher_test_rs_vs_notrs_full_table)), "adj_combined_pvalue_HSO"] < 0.05),
                         sum(fisher_test_rs_vs_notrs_full_table[grep("ALT-Left", rownames(fisher_test_rs_vs_notrs_full_table)), "adj_combined_pvalue_HSO"] > 0.05),
                         sum(fisher_test_rs_vs_notrs_full_table[grep("REF-Middle", rownames(fisher_test_rs_vs_notrs_full_table)), "adj_combined_pvalue_HSO"] < 0.05),
                         sum(fisher_test_rs_vs_notrs_full_table[grep("REF-Middle", rownames(fisher_test_rs_vs_notrs_full_table)), "adj_combined_pvalue_HSO"] > 0.05),
                         sum(fisher_test_rs_vs_notrs_full_table[grep("ALT-Middle", rownames(fisher_test_rs_vs_notrs_full_table)), "adj_combined_pvalue_HSO"] < 0.05),
                         sum(fisher_test_rs_vs_notrs_full_table[grep("ALT-Middle", rownames(fisher_test_rs_vs_notrs_full_table)), "adj_combined_pvalue_HSO"] > 0.05),
                         sum(fisher_test_rs_vs_notrs_full_table[grep("REF-Right", rownames(fisher_test_rs_vs_notrs_full_table)), "adj_combined_pvalue_HSO"] < 0.05),
                         sum(fisher_test_rs_vs_notrs_full_table[grep("REF-Right", rownames(fisher_test_rs_vs_notrs_full_table)), "adj_combined_pvalue_HSO"] > 0.05),
                         sum(fisher_test_rs_vs_notrs_full_table[grep("ALT-Right", rownames(fisher_test_rs_vs_notrs_full_table)), "adj_combined_pvalue_HSO"] < 0.05),
                         sum(fisher_test_rs_vs_notrs_full_table[grep("ALT-Right", rownames(fisher_test_rs_vs_notrs_full_table)), "adj_combined_pvalue_HSO"] > 0.05)),
                       dim = c(2,2,3),
                       dimnames = list(Significant = c("Yes", "No"),
                                       Allele = c("REF", "ALT"),
                                       Location = c("Left", "Middle", "Right")))
# mantelhaen.test(chi_two_array)
```

```{r}
# mantelhaen.test(chi_two_array)
chi_two_df <- data.frame(signif = c(sum(fisher_test_rs_vs_notrs_full_table[grep("REF-Left", rownames(fisher_test_rs_vs_notrs_full_table)), "adj_combined_pvalue_HSO"] < 0.05),
                                    sum(fisher_test_rs_vs_notrs_full_table[grep("REF-Middle", rownames(fisher_test_rs_vs_notrs_full_table)), "adj_combined_pvalue_HSO"] < 0.05),
                                    sum(fisher_test_rs_vs_notrs_full_table[grep("REF-Right", rownames(fisher_test_rs_vs_notrs_full_table)), "adj_combined_pvalue_HSO"] < 0.05)),
                         non_signif = c(sum(fisher_test_rs_vs_notrs_full_table[grep("REF-Left", rownames(fisher_test_rs_vs_notrs_full_table)), "adj_combined_pvalue_HSO"] > 0.05),
                                        sum(fisher_test_rs_vs_notrs_full_table[grep("REF-Middle", rownames(fisher_test_rs_vs_notrs_full_table)), "adj_combined_pvalue_HSO"] > 0.05),
                                        sum(fisher_test_rs_vs_notrs_full_table[grep("REF-Right", rownames(fisher_test_rs_vs_notrs_full_table)), "adj_combined_pvalue_HSO"] > 0.05)))
rownames(chi_two_df) <- c("Left", "Middle", "Right")
chi_two_df$location = rownames(chi_two_df)
chi_two_df_melt <- melt(chi_two_df)
ggplot(chi_two_df_melt, aes(x = location, y = value, fill = location)) + geom_bar(stat = "identity") + 
  facet_wrap(~ variable) +
  ggtitle("Barplot of the number of significant combined adjusted p-value\nfor sample HSO on REF sequences")
```

#### Analysis on ALT sequences for HSO sample

Here is the boxplot of adjusted combined p-values for sample HSO on ALT sequences.

```{r}
chi_two_df <- data.frame(signif = c(sum(fisher_test_rs_vs_notrs_full_table[grep("ALT-Left", rownames(fisher_test_rs_vs_notrs_full_table)), "adj_combined_pvalue_HSO"] < 0.05),
                                    sum(fisher_test_rs_vs_notrs_full_table[grep("ALT-Middle", rownames(fisher_test_rs_vs_notrs_full_table)), "adj_combined_pvalue_HSO"] < 0.05),
                                    sum(fisher_test_rs_vs_notrs_full_table[grep("ALT-Right", rownames(fisher_test_rs_vs_notrs_full_table)), "adj_combined_pvalue_HSO"] < 0.05)),
                         non_signif = c(sum(fisher_test_rs_vs_notrs_full_table[grep("ALT-Left", rownames(fisher_test_rs_vs_notrs_full_table)), "adj_combined_pvalue_HSO"] > 0.05),
                                        sum(fisher_test_rs_vs_notrs_full_table[grep("ALT-Middle", rownames(fisher_test_rs_vs_notrs_full_table)), "adj_combined_pvalue_HSO"] > 0.05),
                                        sum(fisher_test_rs_vs_notrs_full_table[grep("ALT-Right", rownames(fisher_test_rs_vs_notrs_full_table)), "adj_combined_pvalue_HSO"] > 0.05)))
rownames(chi_two_df) <- c("Left", "Middle", "Right")
chi_two_df$location = rownames(chi_two_df)
chi_two_df_melt <- melt(chi_two_df)
ggplot(chi_two_df_melt, aes(x = location, y = value, fill = location)) + geom_bar(stat = "identity") + 
  facet_wrap(~ variable) +
  ggtitle("Barplot of the number of significant combined adjusted p-value\nfor sample HSO on ALT sequences")
```

#### Cochran-Mantel-Haenszel chi-squared test for sample 101

We can have the feeling that Left shifting has more significant p-values. We can validate this by running a Cochran-Mantel-Haenszel chi-squared test. I run it on the following table.

```{r}
ftable(chi_two_array)

chi_two_array <- array(c(3256, 691, 3360, 691,
                         2984, 963, 3038, 1013,
                         2850, 1097, 2872, 1179),
                       dim = c(2,2,3),
                       dimnames = list(Significant = c("Yes", "No"),
                                       Allele = c("REF", "ALT"),
                                       Location = c("Left", "Middle", "Right")))
# mantelhaen.test(chi_two_array)
```

We can look at the odds ratio to have a have an idea of the future result

```{r}
apply(chi_two_array, 3, odds.ratio)
```

The odds ratio are very close. We can think that the significance of the Fisher test on REF or ALT sequence is not dependent to the sequence shifting.

And here is the chi-squared results

```{r}
mantelhaen_results <- mantelhaen.test(chi_two_array)
mantelhaen_results_df <- data.frame(p.value = mantelhaen_results$p.value, odd_ratio = mantelhaen_results$estimate)
rownames(mantelhaen_results_df) <- ""
kable(mantelhaen_results_df)
```

The p-value is not significant, as expected above.
